### 2024-10-18 manuscript.Rmd
revision updates

---
title: 'Cryptobenthic reef assemblages are more distinct across a 90 m depth gradient than 2,500 km of shallow marine habitat in the Hawaiian Archipelago'
author:
  - Mykle L. Hoban:
      institute:
        - himb
      equal_contributor: yes
      correspondence: yes
      email: "mhoban@hawaii.edu"
  - Kaleonani K.C. Hurley:
      institute:
        - himb
      equal_contributor: yes
  - Kerry Reardon:
      institute: noaa
  - Derek J. Skillings:
      institute: 
        - uncg
  - Molly A. Timmers:
      institute:
        - himb
        - natgeo
  - Robert J. Toonen:
      institute: himb
institute:
  - himb: Hawai&#x02BB;i Institute of Marine Biology, University of Hawai&#x02BB;i, Mānoa
  - noaa: National Oceanic and Atmospheric Administration
  - uncg: 'University of North Carolina, Greensboro'
  - natgeo: National Geographic Pristine Seas
bibliography: references/citations.json
csl: references/style.csl
output:
  word_document:
    pandoc_args:
      - "--filter=resources/pandoc-crossref"
      - "--citeproc"
      - "--toc-depth=2"
      - "--lua-filter=resources/scholarly-metadata.lua"
      - "--lua-filter=resources/author-info-blocks.lua"
      - "--lua-filter=resources/abstract-section.lua"
      - "--reference-doc=resources/ref.docx"
      - "--lua-filter=resources/multirefs.lua"
    keep_md: true
tblPrefix: table
figPrefix:
  - "figure"
  - "figures"
chapters: True
chaptersDepth: 1
chapDelim: ""
header-includes:
  - \usepackage{bm}
  - \usepackage{upgreek}
multiref_no_duplicates: true
---
<!--word count: Proceedings B publishes articles up to 12 pages, including figures and tables, with no more than 10 000 words. The word count includes the title page, author names, abstract, figure legends, tables and the reference list. As a rough guide, 10 000 words = 12 pages without additional figures; 850 words = 1 page and 2-4 small figures = 1 page. For example, an acceptable paper could be made up of 9 000 words and 4 small figures or 8500 words, 2 large figures and 4 small figures.-->
### Keywords {-}
<div custom-style="mono">Mesophotic coral ecosystems, community structure, depth gradient, environmental drivers, Hawaii, Autonomous Reef Monitoring Structures</div>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  fig.path="output/figures/fig-",
  dev="svg",
  dpi=300,
  message=FALSE,
  warning=FALSE,
  cache=TRUE,
  cache.path = "cache/",
  cache.comments=FALSE
)

library(tidyverse)
library(pander)
library(fs)
library(here)
library(english)
library(phyloseq) 
library(vegan)
library(ggrepel)
library(here)
library(worrms)
library(ggtext)

# can't get BiodiversityR to work on new mac, but that's ok because
# all we need is the CAPdiscrim function, which miraculously works on its own!
# retrieved from: https://github.com/cran/BiodiversityR
source(here("lib","CAPdiscrim.R"))

```

```{r init, include=FALSE}
# source utility functions
source("crabs.R")

# setup community matrices and data summaries
if (!exists("crabs_done")) {
  crabs_done <<- FALSE
}
if (!crabs_done) {
  cc <<- setup_crabs()
  crabs_done <<- TRUE
}

# some utility funtions

# do an english comma-separated list
format_list <- function(l) {
  l[length(l)] <- str_c("and ",l[length(l)])
  if (length(l) > 2) { 
    str_c(l,collapse=", ")
  } else {
    str_c(l,collapse=" ")
  }
}

# format a range of numbers
format_range <- function(vals) {
  f <- range(vals)
  str_glue("{f[1]}--{f[2]}")
}

# quickie to just give us a depth range
dr <- function(ps) {
  ps %>%
    sample_tibble() %>%
    pull(depth) %>%
    format_range()
}

# basic reusable ggplot theme
plotz_theme <- function() {
  theme_bw() + 
    theme(
      legend.key = element_blank(),
      panel.grid = element_blank()
    )
}

# abstract method to format species name
format_species <- function(...) UseMethod("format_species")

# format species in a list
format_species.list <- function(spp,authority=TRUE,...) {
  if (spp$found) {
    a <- ifelse(authority,str_glue(" {spp$authority}"),"")
    str_glue("*{spp$species}*{a}")
  } else {
    s <- str_split(spp$species," ")[[1]]
    str_trim(str_glue("*{s[1]}* {s[2]}"))
  }
}

# format species in a character vector
format_species.character <- function(species,found,authority,auth=TRUE,...) {
  spp <- map2_chr(species,found,~{
    .y <- ifelse(is.na(.y),FALSE,.y)
    s <- list(species=.x,found=.y,authority="")
    format_species(s,FALSE)
  })
  if (auth) {
    authority <- if_else(!is.na(authority),str_c(" ",authority),"")
    spp <- str_c(spp,authority)
  }
  return(spp)
}

# format species in a data frame
format_species.data.frame <- function(spp,authority=TRUE,...) {
  spp %>%
    pmap_chr(~{
      row <- list(...)
      s <- row[c('species','found','authority','otu')]
      format_species(s,authority)
    })
}

# format percents and numbers
pc <- function(x,a=0.01) scales::percent_format(accuracy = a)(x)
num <- function(x,a=0.01) scales::label_comma(accuracy=a)(x)

# format wilcoxon test results
w_test <- function(test,plain=FALSE) {
  if (plain) {
    str_glue("*W* = {num(test$statistic,0.01)}, {pval(test$p.value)}")
  } else {
    str_glue("(Wilcoxon rank sum test: *W* = {num(test$statistic,0.01)}, {pval(test$p.value)})")
  }
}

# format kruskal-wallis test results
k_test <- function(test,plain=FALSE) {
  if (plain) {
    str_glue("*K-W $\\upchi^2$* = {num(test$statistic,0.01)}, DF = {num(test$parameter['df'],0.01)}, {pval(test$p.value)}")
  } else {
    str_glue("Kruskal-Wallis rank sum test: *K-W $\\upchi^2$* = {num(test$statistic,0.01)}, DF = {num(test$parameter['df'],0.01)}, {pval(test$p.value)}")
  }
}

# format t-test results
t_test <- function(test,plain=FALSE,paren=FALSE) {
  if (plain) {
    str_glue("*t* = {num(test$statistic,0.01)}, DF = {num(test$parameter['df'],0.01)}, {pval(test$p.value)}")
  } else {
    str_glue("{ifelse(paren,'(','')}{test$method}: *t* = {num(test$statistic,0.01)}, DF = {num(test$parameter['df'],0.01)}, {pval(test$p.value)}{ifelse(paren,')','')}")
  }
}

# format f-test results
f_test <- function(test,plain=FALSE,paren=FALSE) {
  if (plain) {
    str_glue("*F* = {num(test$statistic,0.01)}, DF = $\\frac{{{test$parameter[1]}}}{{{test$parameter[2]}}}$, {pval(test$p.value)}")
  } else {
    str_glue("{ifelse(paren,'(','')}F-test: *F* = {num(test$statistic,0.01)}, DF = $\\frac{{{test$parameter[1]}}}{{{test$parameter[2]}}}$, {pval(test$p.value)}{ifelse(paren,')','')}")
  }
}

# format a p-value nicely
pval <- function(p) {
  case_when(
    p < 0.001 ~ "*p* \\< 0.001",
    p < 0.01 ~ "*p* \\< 0.01",
    TRUE ~ as.character(str_glue("*p* = {num(p,0.01)}"))
  )
}

# format permdisp test results
permdisp <- function(bd,accuracy=0.01,perm=999) {
  a <- anova(bd,permutations=perm) %>%
    broom::tidy() %>%
    filter(term == "Groups")
  with(a,str_glue("DF = {num(df,accuracy)}, SS = {num(sumsq,accuracy)}, F = {num(statistic,accuracy)}, {pval(p.value)}"))
}

# format permanova test results
permanova <- function(modl,accuracy=0.01,stat="pseudo-F",by=NULL,sentence=TRUE) {
  has_r2 <- TRUE
  if (is(modl,"dbrda")) {
    modl <- anova(modl,by=by)
    has_r2 <- FALSE
  }
  m <- suppressWarnings(broom::tidy(modl)) %>%
    filter(!term %in% c("Total","Residual"))
  if (nrow(m) > 1) {
    m %>%
      pmap_chr(~{
        with(list(...), {
          if (has_r2) {
            r2 <- str_glue(" $\\mathrm{{R}}^2$ = {num(R2,0.01)},") 
          } else {
            r2 <- ""
          }
          if (sentence) {
            term <- str_replace_all(term,"_"," ")
          }
          str_glue("{term}: DF = {num(df,accuracy)}, SS = {num(SumOfSqs,accuracy)},{r2} {stat} = {num(statistic,accuracy)}, {pval(p.value)}")
        })
      }) %>%
      str_c(collapse = "; ")
  } else {
    with(m %>% slice(1),{
      if (has_r2) {
        r2 <- str_glue(" $\\mathrm{{R}}^2$ = {num(R2,accuracy)},") 
      } else {
        r2 <- ""
      }
      if (sentence) {
        term <- str_replace_all(term,"_"," ")
      }
      str_glue("{term}: DF = {num(df,accuracy)}, SS = {num(SumOfSqs,accuracy)},{r2} {stat} = {num(statistic,accuracy)}, {pval(p.value)}")
      # str_glue("DF = {num(df,accuracy)}, SS = {num(SumOfSqs,accuracy)}, $\\mathrm{{R}}^2$ = {num(R2,accuracy)}, pseudo-F = {num(statistic,accuracy)}, {pval(p.value)}")
    }) 
  }
}

# format a top-n list of species
top_list <- function(species,n,units,len=4,fig=NULL) {
  s1 <- map_chr(species,format_species)
  s2 <- map2_chr(n,units,~{
    if (is.null(fig)) {
      str_glue("({.x} individuals from {.y} units)")
    } else {
      str_glue("({.x} individuals from {.y} units; [@fig:{fig}[]{{{{}}}}{{.y}}])")
    }
  })
  if (!is.null(fig)) {
    s2 <- map2_chr(s2,letters[seq(length(s2))],~str_glue(.x))
  }
  ss <- str_c(s1," ",s2)
  ss[1:(len-1)] %>%
    str_c(collapse=", ") %>%
    str_c(", and ",ss[len])
}

# format a markdown grid table from a data frame or file
print_table <- function(tbl,id="",caption="",hdr_case="sentence",hide=character(0),accuracy = 0.01,format_numbers=TRUE, cell_width = Inf) {
  if (!grepl("^: ",caption)) {
    caption <- paste0(": ",caption)
  }

  if (inherits(tbl,"character")) {
    if (id == "") {
      id <- sub(pattern = "(.*?)\\..*$", replacement = "\\1", basename(file))
    }
    tbl <- read.csv(tbl,check.names=FALSE,stringsAsFactors=FALSE)
  } else if (inherits(tbl,"data.frame")) {
    if (id == "") stop("please specify table id")
  } else {
    stop("'tbl' must be a character or a data frame")
  }

  tbl <- tbl %>%
    select(-any_of(hide)) %>%
    mutate(
      across(
        where(~format_numbers & is.numeric(.x)),
        ~num(.x,accuracy)
      )
    )

  # can also use rename_with(str_to_sentence)
  names(tbl) <- switch(
    str_to_lower(hdr_case),
    sentence = str_to_sentence(names(tbl)),
    title = str_to_title(names(tbl)),
    upper = str_to_upper(names(tbl)),
    lower = str_to_lower(names(tbl)),
    names(tbl)
  )

  caption <- sprintf("%s {#tbl:%s}",caption,id)

  table_str <- pander::pandoc.table.return(
    tbl,
    keep.line.breaks=TRUE,
    caption=NULL,
    style="grid",
    split.cells=cell_width,
    split.tables=Inf,
    justify='left',
    missing=''
  )
  table_str <- table_str[table_str != ""]
  cat(paste0(table_str,collapse="\n"),"\n\n",caption,"\n\n",sep="")
}

# get average distance from a centroid
centroid_dist <- function(data,mean=TRUE) {
  pts <- t(as.matrix(data))
  centers <- matrix(apply(pts,1,mean),nrow=2)
  dists <- sqrt(apply(centers, 2, function(center) {
    colSums((pts - center)^2)
  }))
  if (mean) {
    return(mean(dists))
  } else {
    return(c(dists))
  }
}

# get alpha diversity statistics
all_alpha <- alpha_diversity(cc$crabs_untransformed)
deep_alpha <- alpha_diversity(cc$crabs_deep_untransformed)
shallow_alpha <- alpha_diversity(cc$crabs_shallow_untransformed)

# color palette for islands
island_pal <- c(
  "Hawaii Island" = "#5E4FA2",
  "Hawai‘i Island" = "#5E4FA2",
  "Maui" = "#3288BD",
  "Oahu" = "#66C2A5",
  "O‘ahu" = "#66C2A5",
  "Kauai" = "#ABDDA4",
  "Kaua‘i" = "#ABDDA4",

  "French Frigate Shoals" = "#FDAE61",
  "Lalo" = "#FDAE61",
  "Lalo (French Frigate Shoals)" = "#FDAE61",
  "Lisianski" = "#F46D43",
  "Kapou" = "#F46D43",
  "Kapou (Lisianski)" = "#F46D43",
  "Pearl and Hermes Reef" = "#D53E4F",
  "Manawai" = "#D53E4F",
  "Manawai (Pearl and Hermes Reef)" = "#D53E4F",
  "Kure Atoll" = "#9E0142",
  "Hōlanikū" = "#9E0142",
  "Hōlanikū (Kure Atoll)" = "#9E0142"
)

# color palette for island groups
island_group_pal <- c(
  "main" = "#ffb400",
  "Main" = "#ffb400",
  "Northwest" = "#9080ff",
  "northwest" = "#9080ff"
)

# colors for sample group / depth zone
sample_depth_pal <- c(
  "shallow" = "#ffb400",
  "Shallow" = "#ffb400",
  "Shallow (MHI)" = "#ffb400",
  "shallow_main" = "#ffb400",
  "Shallow (NWHI)" = "#9080ff",
  "shallow_nwhi" = "#9080ff",
  "30 m" = "#d87939",
  "30m" = "#d87939",
  "60 m" = "#04c3c8",
  "60m" = "#04c3c8",
  "90 m" = "#04738d",
  "90m" = "#04738d",
  "deep" = "#04738d",
  "Deep" = "#04738d",
  "O‘ahu (shallow)" = "#e5ab52",
  "O‘ahu 30 m" = "#d87939",
  "O‘ahu 30m" = "#d87939",
  "O‘ahu 60 m" = "#04c3c8",
  "O‘ahu 60m" = "#04c3c8",
  "O‘ahu 90 m" = "#04738d",
  "O‘ahu 90m" = "#04738d",
  "O‘ahu deep" = "#04738d",
  "O‘ahu Deep" = "#04738d"
)

# number of permutations for adonis tests
PERM <<- 999
# total number of brachyuran species in Hawai‘i
hawaiian_crabs <- 284
# load model terms table from file
model_terms <- read_csv(here("data","tables","environmental_variables.csv"))
```

```{r taxonomy-validation, include=FALSE, dependson="init"}
# validate taxonomy using worms / get authorities
# worms API call results are cached
crab_taxonomy <- cc$crabs %>%
  taxa_tibble() %>%
  pull(species) %>%
  # split into subsets of 50 taxa
  split(ceiling(seq_along(.)/50)) %>%
  map(\(sp) {
    # validate taxa using fuzzy matching
    wm_records_taxamatch(sp) %>%
      # smash results into a single tibble
      map2(sp,~.x %>% mutate(query = .y)) %>%
      list_rbind()
  }) %>% 
  list_rbind()

# max validated taxonomy table
crab_taxonomy_validated <- cc$crabs %>%
  taxa_tibble() %>%
  left_join(crab_taxonomy,by=c("species" = "query")) %>%
  mutate(
    found = rank == "Species",
    # substitute valid name if ours isn't
    species = case_when(
      valid_name != scientificname ~ valid_name,
      .default = species
    ),
    authority = case_when(
      valid_name != scientificname ~ valid_authority,
      .default = authority
    )
  ) %>%
  mutate(display_species = format_species(.))
```

```{r summaries, include=FALSE, dependson="init"}
# generate sample summaries. code block is cached 
# so worms API calls don't happen every time
all_summary <- sample_summary(cc$crabs_untransformed,4,crab_taxonomy_validated)
deep_summary <- sample_summary(cc$crabs_deep_untransformed,4,crab_taxonomy_validated)
```

<!-- 
the abstract-section.lua filter allows us to put our abstract down here 
which allows us to include computed values that couldn't easily go in the yaml header
-->
# Abstract {-}

An inability to consistently resolve the environmental and anthropogenic factors that shape coral reefs impacts our understanding of how those factors interact and how coral reef communities may respond to change. Logistical challenges have additionally limited the taxonomic resolution and depth ranges investigated, with the focus largely on fishes and corals in the shallowest ~30 m of coral reefs. Here, we examine the effects of a suite of environmental, ecological, and anthropogenic factors on the assemblage structure of brachyuran crabs recovered from Autonomous Reef Monitoring Structures on shallow reefs (`r dr(cc$crabs_shallow_untransformed)` m) across the ~2,500 km span of the Hawaiian Archipelago. We compare these assemblages with those sampled along a depth gradient (`r format_range(range(cc$crab_data$depth[cc$crab_data$island == "Oahu"]))` m) on O&#x02BB;ahu. We recovered `r num(sum(sample_sums(cc$crabs_untransformed)),1)` individuals representing `r sum(unlist(all_summary$types_breakdown))` taxa from `r all_summary$taxonomy_breakdown$families` families. Shallow assemblages were significantly different between the Main and Northwest Hawaiian Islands and significantly associated with human impacts, chlorophyll-A, sea-surface temperature, slope, potential larval immigration, and depth. However, we found more variation among mesophotic samples from O&#x02BB;ahu (30--90 m) than among shallow samples across the entire archipelago, suggesting that 90 m of depth is a stronger determinant of brachyuran assemblage structure than the latitudinal, environmental, or anthropogenic gradients across the Hawaiian Islands.

# Introduction {-}

Coral reefs are among the most diverse habitats in the world, and they face increasing threats of degradation, including climate-driven bleaching and phase shifts to alternate ecological regimes [@grahamManagingResilienceReverse2013; @norstromAlternativeStatesCoral2009; @pandolfiGlobalTrajectoriesLongterm2003]. In the face of these threats, understanding drivers of metazoan community structure in coral reef ecosystems is critically important [@hughesRisingChallengeSustaining2010; @pageSeekingResistanceCoral2019], particularly as the search for criteria to prioritize climate refugia such as upwelling zones or mesophotic reefs intensifies. While the structure and composition of coral reef communities is influenced by a suite of environmental and anthropogenic factors, the nature and makeup of those factors remain poorly resolved [@hughesCommunityStructureDiversity1989; @smithEffectsTopBottom2010; @brandlCoralReefEcosystem2019]. One aspect of coral reef community structure that is thought to be important but remains under-investigated is depth. Most studies to date have focused on the upper ~30 m of the water column, yet coral reefs and associated communities are known to extend beyond 150 m [@pyleExploringDeepCoral1996; @pyleMesophoticCoralEcosystems2019], with photosymbiotic corals reported to 170 m [@rouzeSymbioticAssociationsDeepest2021]. As much as two thirds of coral reef ecosystems lie in this under-explored zone [@hindersteinThemeSectionMesophotic2010]. These habitats, known as mesophotic coral ecosystems (MCEs), are typically defined as those extending from ~30--150 m depth [@hindersteinThemeSectionMesophotic2010; @kahngCommunityEcologyMesophotic2010]. They have been hypothesized as possible refugia from climate change impacts on shallow reefs [@glynnCoralReefBleaching1996; @bongaertsAssessingDeepReef2010; @smithDepthRefugiumCatastrophic2014; @holsteinFertileFathomsDeep2015; @muirSpeciesIdentityDepth2017], although a growing number of studies have found evidence that mesophotic reefs are unique and uniquely threatened, challenging this notion [@diazLightTemperatureDrive2023; @diazMesophoticCoralBleaching2023; @hobanPlumbingDepthsEnvironmental2023; @hurleyAssessmentShallowMesophotic2016; @kaneHighLevelsMesophotic2014; @pinheiroDeepReefFishes2019; @rochaMesophoticCoralEcosystems2018; @montgomeryCommunitySimilaritySpecies2021].

The Hawaiian Archipelago, stretching across ~2,500 km in the North Pacific Ocean, provides a natural laboratory to examine drivers of metazoan community structure. This chain of islands comprises two regions, the inhabited high islands of the main Hawaiian Islands (MHI) and the uninhabited low islands, atolls, and banks of the Northwestern Hawaiian Islands (NWHI) which reside within one of the world's largest marine protected areas---the Papahānaumokuākea Marine National Monument. The island chain is subject to considerable anthropogenic and physical gradients driving differences in temperature, coral cover, human population, and anthropogenic impact, among others, all of which influence coral reef communities [@selkoeMapHumanImpacts2009]. Nearshore drop-offs along the island chain also allow access to deeper reef habitat, enabling the examination of depth influences on community structure.

To date, the majority of studies that have examined community structure across the Hawaiian archipelago and along a depth gradient have focused on conspicuous taxa such as corals and fishes [e.g., @benfieldComparisonCoralReef2008; @caballero-aragonWaveExposureTemperature2023; @kaneTrophicDesignationLive2017; @langeWaveExposureShapes2021; @vanlierHabitatConnectivityComplexity2018]. However, these taxa make up a minority of coral reef biodiversity and thus may not be indicative of community patterns overall [@fautinOverviewMarineBiodiversity2010]. Furthermore, studies on these taxa have also been criticized for confounding multiple potentially covarying factors (such as habitat area and/or complexity) across the measured gradients [e.g., @bellwoodRegionalscaleAssemblyRules2001; @grahamEffectsMarineReserve2003; @connollyCommunityStructureCorals2005; @ewersConfoundingFactorsDetection2006; @fletcherRoleHabitatArea2007].

Here, we used autonomous reef monitoring structures (ARMS)---standardized collection devices designed to mimic reef structural complexity---to sample cryptic communities while controlling for habitat [@zimmermanArtificialReefMatrix2004; @brainardAutonomousReefMonitoring2009]. Such standardization can reduce variability in recovered communities that may arise due to local variation in habitat composition and availability [@ransomeImportanceStandardizationBiodiversity2017]. We targeted brachyuran crabs, a ubiquitous group often used as an exemplar to assess coral reef biodiversity [@knowltonCoralReefBiodiversity2010; @plaisanceDiversityCoralReefs2011; @servisCharacterizingCoralReef2020; @plaisanceReefassociatedCrustaceanFauna2009] and comprising a notable proportion of the coral reef cryptofauna [284 species in Hawai&#x02BB;i, @castroCatalogAnomuranBrachyuran2011]. We used assemblages recovered from ARMS units to examine drivers of community composition across the Hawaiian Archipelago, compared crab communities among islands and regions (MHI vs. NWHI), and examined the effects of a suite of possible environmental and anthropogenic drivers of community structure to test hypotheses about their relative importance.

Much of our understanding of biotic zonation and community structure in the ocean originates with classic studies from the intertidal [e.g., @connellInfluenceInterspecificCompetition1961; @connellEffectsCompetitionPredation1961; @paineIntertidalCommunityStructure1974]. While there are no shortages of exceptions, a long history of studies point towards a generalization that the lower bathymetric limits are generally regulated by ecological interactions whereas upper limits of intertidal communities are more often determined by abiotic factors [@stephenson1972life; @mengeRockyIntertidalCommunities2001; @bertnessTestingRelativeContribution1999; @dennyPhysicalProcessesThat2001]. However, most of these models describe zonation at specific localities rather than community variation over larger spatial scales [but see @birdTideWaveMarks2013; @bishop-taylorTidesModellingElevation2019] and may not apply to coral reefs, which are subject to more complex community dynamics [@abramsRecruitmentLotteriesCoexistence1984; @bussCompetitiveNetworksNontransitive1979; @hixonPredationPreyRefuges1993; @hubbellUnifiedNeutralTheory2011; @moraEcologyFishesCoral2015; @saleStockrecruitRelationshipsRegional1982; @sheppardBiologyCoralReefs2017]. While there is little doubt that ecological interactions are important on coral reefs, the suite of factors and which play the most important role in structuring coral reef communities remains a subject of considerable debate [@bellwoodRegionalscaleAssemblyRules2001; @connollyUnifiedModelExplains2017; @hubbellUnifiedNeutralTheory2011; @jacksonAdaptationDiversityReef1991; @mumbyCoralReefHabitats2008; @saleMaintenanceHighDiversity1977].

Given the strong reported differences between deep and shallow reefs across taxa and the paucity of research on drivers of cryptic coral reef community variation, we sought to answer the following four research questions: (i) How do shallow reef brachyuran crab assemblages vary along the natural and anthropogenic gradients of the Hawaiian Archipelago? (ii) Do shallow reef brachyuran crab assemblages differ between the Main and Northwestern Hawaiian Islands? (iii) What environmental and anthropogenic factors drive crab assemblage structure in shallow coral reefs across the Hawaiian Archipelago? (iv) What are the relative contributions of depth (shallow vs. mesophotic) and island region to crab community variation across the Hawaiian Archipelago?

# Materials and Methods {-}

## Sample Collection {-}

Brachyuran crab assemblages were sampled from benthic coral reef habitat in the Hawaiian Archipelago using Autonomous Reef Monitoring Structures (ARMS) collected by NOAA's Ecosystem Science Division as part of its Pacific Reef Assessment and Monitoring Program. ARMS are standardized collection devices designed to mimic coral reef structural complexity to attract settlement of reef cryptofauna [@zimmermanArtificialReefMatrix2004; @brainardAutonomousReefMonitoring2009]. They are constructed from 10 gray type 1 PVC plates (23 cm $\times$ 23 cm) stacked in an alternating series of open and semi-enclosed layers ([@fig:fig_arms]). The stack is fastened to a larger base plate (35 cm $\times$ 45 cm) which can be attached to the benthos upon deployment [@brainardAutonomousReefMonitoring2009; @hurleyAssessmentShallowMesophotic2016; @knowltonCoralReefBiodiversity2010; @zimmermanArtificialReefMatrix2004].

![Autonomous reef monitoring structures (ARMS) deployed in the Hawaiian Archipelago.](data/images/arms_example.png){#fig:fig_arms}

<br><br>

A total of 89 ARMS units were deployed over the course of six years (2009--2016) in shallow (`r dr(cc$crabs_shallow_untransformed)` m depth) coral reef habitat in both the Main (MHI) and Northwest Hawaiian Islands (NWHI) ([@fig:fig_map; @tbl:arms_table]). Only units deployed on outer reefs were included in this study. To explore the effects of depth, we also present data from an earlier study by Hurley et al. [-@hurleyAssessmentShallowMesophotic2016], in which 27 units were deployed off O&#x02BB;ahu from 2009--2012 in `r format_range(cc$crab_data %>% sample_tibble() %>% filter(island == "Oahu") %>% pull(depth))` m depth; nine shallow units were deployed at 12 m, and six each at depths of 30 m, 60 m, and 90 m ([@tbl:arms_table]). Shallow units were deployed according to NOAA-RAMP methodology, whereas units $\geq$ 30 m were deployed and recovered using technical mixed gas closed-circuit rebreather diving [see @hurleyAssessmentShallowMesophotic2016]. Except when discussing specific depth zones, we will refer to any units deployed at `r dr(cc$crabs_shallow_untransformed)` m as "shallow" and those deployed $\geq$ 30 m as "deep".

```{r fig-map, fig.cap='Map of ARMS deployment sites across the Hawaiian Archipelago (Papahānaumokuākea Marine National Monument (NWHI) is shaded in light blue).', results="asis", fig.width=9, fig.height=6, dependson="init"}
library(sf)
library(rnaturalearth)
library(rnaturalearthhires)

crab_sites <- cc$crabs_untransformed %>%
  sample_tibble() %>%
  distinct(hawaiian_name,island,island_group,lat,lon,.keep_all = FALSE) %>%
  mutate(
    island = case_when(
      str_remove_all(hawaiian_name,'[^a-zA-Z ]') != island ~ as.character(str_glue("{hawaiian_name} ({island})")),
      TRUE ~ as.character(hawaiian_name)
    ),
    island = fct_reorder(island,-lat)
  ) %>%
  st_as_sf(coords=c("lon","lat"),crs=4326, remove=F) #%>%

hawaii <- ne_states("united states of america",returnclass = "sf") %>%
  filter(name == "Hawaii")

trans_box <- function(b,crs) {
  box_before <- st_sfc(
    st_point(b[c(1,2)]),
    st_point(b[c(3,4)]),
    crs=4326
  )
  box_before %>%
    st_transform(crs) %>%
    st_bbox()
}

shift_box <- function(b) {
  if (b['xmin'] < 0) 
    b['xmin'] <- b['xmin'] + 360
  if (b['xmax'] < 0) 
    b['xmax'] <- b['xmax'] + 360
  return(b)
}

pmnm <- st_read(here("data","gis","hi_noaa_nwhi_papahanaumokuakea.shp"),quiet = TRUE) %>%
  st_shift_longitude()

pbox <- st_bbox(pmnm)
cbox <- st_bbox(st_buffer(crab_sites,units::as_units(50,"km")) %>% st_shift_longitude())
hbox <- hawaii %>%
  st_shift_longitude() %>%
  st_bbox()

box_before <- c(
  xmin = min(c(pbox['xmin'],hbox['xmin'],hbox['xmin'])),
  ymin = min(c(pbox['ymin'],cbox['ymin'],hbox['ymin'])),
  xmax = max(c(pbox['xmax'],cbox['xmax'],hbox['xmax'])) + 1.5,
  ymax = max(c(pbox['ymax'],cbox['ymax'],hbox['ymax']))
)

transformer <- str_glue("+proj=leac +lat_1={box_before[2]} +lat_2={box_before[4]} +lon_0={mean(box_before[c(1,3)])}")

box_after <- box_before %>%
  trans_box(transformer)

pmnm <- pmnm %>%
  st_transform(transformer)

ggplot() +
  geom_sf(data=hawaii %>% st_transform(transformer), color="black", fill="#A3866A", linewidth=0.4) +
  geom_sf(data=pmnm,color=NA,fill="#6ed8f6",alpha=0.2) +
  geom_sf(data=crab_sites %>% st_transform(transformer),aes(fill=island),size=3,shape=21,color="black") +
  scale_fill_manual(values=island_pal,name="Island") +
  coord_sf(xlim=box_after[c(1,3)],ylim=box_after[c(2,4)]) +
  scale_x_continuous(breaks = -seq(180, 155, -5)) +
  theme_minimal() +
  xlab("Longitude") +
  ylab("Latitude") +
  theme(legend.text = element_text(size=10),legend.title = element_text(size=12))
cat("{#fig:fig_map}")
```

<br><br>

```{r arms-table,results="asis", dependson="init"}
arms_schedule <- cc$crabs_untransformed %>%
  sample_tibble() %>%
  # separate_wider_regex(sample,patterns=c(recovery_year="^[0-9]+","_",sample=".*"))  %>%
  # select(sample,everything()) %>%
  unite("range",deployment_year,recovery_year,sep="--") %>%
  mutate(
    island = case_when(
      str_remove_all(hawaiian_name,'[^a-zA-Z ]') != island ~ as.character(str_glue("{hawaiian_name} ({island})")),
      TRUE ~ as.character(hawaiian_name)
    ),
    island = case_when(
      island == "O‘ahu" ~ str_glue("{hawaiian_name} ({shallow_deep})"),
      TRUE ~ island
    ),
    island = fct_reorder(island,-lat)
  ) %>%
  count(island,range) %>%
  pivot_wider(names_from="range",values_from="n")  %>%
  select(island,order(colnames(.))) %>%
  mutate(across(-island,~replace_na(.x,0)))
print_table(arms_schedule,"arms_table","ARMS unit deployment schedule by island (number of units)",accuracy=1) 
```

<br><br>

ARMS units were retrieved by encapsulating them in-situ in a 100 μm mesh-lined crate to prevent escape of motile organisms. Once returned to the surface, they were placed in containers of aerated, filtered seawater, crates were removed, and units were systematically disassembled plate-by-plate [see @lerayDNABarcodingMetabarcoding2015 for processing details]. All adult brachyuran crabs ($\geq$ 2 mm) were photographed, identified, and preserved in 95% ethanol. Specimens were initially identified to family or genus using Poore & Ahyong [-@pooreMarineDecapodCrustacea2023] and additionally to species where possible according to expert advice from Scott Godwin (former curator of invertebrate zoology, Bishop Museum) and Gustav Paulay (curator of invertebrate zoology, Florida Museum of Natural History). Where species-level identifications were not possible, taxa were collapsed to family or genus (depending on individual identification). We recognize that the species-level taxonomy of Indo-Pacific brachyurans remains poorly resolved, particularly for abundant groups like the Chlorodiellinae [@lasleyjrIntegrativeSystematicRevision2014; @lasleyjrPhylogeneticRelationshipsUbiquitous2015], and some "species" are known to comprise complexes [e.g. *Perinia tumida*, @servisCharacterizingCoralReef2020; @plaisanceReefassociatedCrustaceanFauna2009]. However, the prevalence of species complexes and cryptic diversity suggests that our assignments are likely to under- rather than overrepresent the true number of taxa. Thus, any significant patterns detected likely persist in spite of rather than because of cryptic diversity.

## Data analysis {-}

Unless otherwise specified, all analyses were performed using `R` version `r R.version$major`.`r R.version$minor` [@rcoreteamLanguageEnvironmentStatistical2021]. Community matrices were loaded and encapsulated into data objects using the R package `phyloseq` [@mcmurdiePhyloseqPackageReproducible2013] and multivariate community analyses were performed using the package `vegan` [@oksanenVeganCommunityEcology2022] on Bray-Curtis dissimilarities of Hellinger-transformed species abundance matrices [@legendreEcologicallyMeaningfulTransformations2001; @raoReviewCanonicalCoordinates1995].

To examine patterns of crab diversity across the islands, we used `estimate_richness` to calculate species richness and effective Simpson diversity (effective Simpson = $\frac{1}{\sum_{i=1}^{R}p_i^2}$, where R is the total number of species, and $\mathrm{p}_i$ is the proportional abundance of each species). The effective Simpson index (the inverse of the "classic" Simpson entropy), which has units of species, is easier to interpret in an ecological context than an entropic non-linear index [@jostEntropyDiversity2006; @tuomistoUpdatedConsumerGuide2012]. We used nonparametric Kruskal-Wallis ranked sum tests to compare diversity between shallow and mesophotic samples and between island groups (MHI vs. NWHI). Island group comparisons were performed both including and excluding mesophotic samples.

We used distance-based redundancy analysis (dbRDA; a form of multivariate multiple regression) with `dbrda` to examine the relative contribution of environmental variables ([@tbl:env_table]) to shallow crab community variation across the islands, conditioning the recovery year as a random effect. From a model incorporating all potential drivers, we sequentially removed variables with the largest variance inflation factors (VIF; calculated using `vif.cca`) until all values were <3, indicating variables were sufficiently uncorrelated. Model significance was assessed using permutation tests with `r num(PERM,1)` permutations with the function `anova`. Continuous variables were scaled to unit variance. dbRDA analyses were conducted only for communities sampled from shallow (\< 30 m) ARMS units, since detailed driver variables were not available for deep ($\geq$ 30 m) samples.

We assessed variation among groups using `betadisper` and `adonis2`. These functions implement analysis of multivariate homogeneity of group dispersions (PERMDISP) and multivariate permutational analysis of variance (PERMANOVA) respectively. PERMDISP is used to assess variability within groups and PERMANOVA is used to assess differences among groups. Significance was assessed with permutation tests in both cases. We used PERMANOVA to make the following comparisons: island group (MHI vs NWHI, shallow samples only), island (shallow samples only), depth zone only (all samples), depth zone only (O&#x02BB;ahu samples), island group and depth zone (all samples). Variation among island groups and sampling depths was further examined and visualized using canonical analysis of principal coordinates (CAP) based on discriminant analysis [@andersonCanonicalAnalysisPrincipal2003] as implemented in the `CAPdiscrim` function from `BiodiversityR` [@kindtTreeDiversityAnalysis2005]. To explore differences in community variability among groups (shallow/deep, O&#x02BB;ahu/all shallow samples), we compared Euclidean distances among samples in the CAP ordination using Welch two-sample t-tests, assuming unequal variances.

## Environmental variables {-}

Potential drivers of brachyuran crab community structure were assembled from previous studies [@selkoeDNACoralReef2016; @halpernRecentPaceChange2019], NOAA's National Coral Reef Monitoring Program (NCRMP), the Pacific Islands Ocean Observing System (PacIOOS), and MODIS satellite data products compiled by NASA's Physical Oceanography/Ocean Biology Distributed Active Archive Centers [@nasagoddardspaceflightcenterModerateresolutionImagingSpectroradiometer2022; @nasaobpgMODISAquaGlobal2020]. Variables were either intrinsic (site depth, latitude, longitude, island slope) or derived from remote sensing and/or modeling (chlorophyll-A, sea-surface temperature, coral cover, potential larval immigration, cumulative human impacts) ([@tbl:env_table; @tbl:arms_metadata_table]). Sea-surface temperature (SST) and chlorophyll-A values were calculated by averaging monthly means across the duration of the soaking period for each deployment using AQUA MODIS 4 km-resolution gridded datasets. The cumulative human impacts index, a dimensionless numerical value representing spatially-explicit cumulative anthropogenic impacts on various marine ecosystems, is based on a synthesis of a combination of "high resolution, annual data on the intensity of 14 human stressors and their impact on 21 marine ecosystems over 11 years (2003--2013)" [@halpernGlobalMapHuman2008; @halpernRecentPaceChange2019]. The index of potential larval immigration was imported from [@selkoeDNACoralReef2016] and represents the potential of a given island to act as a destination for larval settlement. It is an "estimate of in-closeness centrality" from a biophysical model simulating larval dispersal in the Hawaiian Islands [@selkoeDNACoralReef2016 supp. mat.]. For details regarding data import, preparation, and synthesis, see Supplemental material. Due to logistical limitations, environmental data were collected only for shallow samples (\< 30 m).

```{r env-table,results="asis", dependson="init"}
print_table(model_terms,"env_table","Potential environmental drivers of crab community structure across the Hawaiian Archipelago",hide="variable")
```

<br><br>

# Results {-}

## Diversity summary {-}

### All ARMS {-}

<!-- TODO: update number of different types to reflect family/genus/species level rather than species/morphospecies level -->
The `r nsamples(cc$crabs_untransformed)` ARMS units deployed across the Hawaiian Islands (in `r format_range(cc$crab_data$depth)` m depth) yielded a total of `r num(sum(sample_sums(cc$crabs_untransformed)),1)` adult crabs. Units contained `r num(all_summary$abundance_mean,0.1)` $\pm$ `r num(all_summary$abundance_sd,0.1)` (mean $\pm$ SD) individuals, and abundance ranged from `r format_range(all_summary$abundance_range)`. Crabs were identified to `r sum(unlist(all_summary$types_breakdown))` distinct taxa (`r all_summary$types_breakdown$species` species-level identifications, `r all_summary$types_breakdown$genus` genus, `r all_summary$types_breakdown$family` family) encompassing `r all_summary$taxonomy_breakdown$genera` genera in `r all_summary$taxonomy_breakdown$families` families, which accounted for ~`r pc(all_summary$types_breakdown$species / hawaiian_crabs,0.1)` of the known crab diversity reported from Hawaiian waters for species-level IDs [@castroCatalogAnomuranBrachyuran2011]. The four most abundant species were: `r with(all_summary,top_list(top,top_n,top_units,fig="fig_crabs"))`. There were `r all_summary$singleton_count` singleton taxa (taxa occurring in a single unit across all sites), accounting for ~`r pc(all_summary$singleton_count/ntaxa(cc$crabs_untransformed),0.1)` of total recovered crab taxa. For a breakdown of specific crab taxa recovered by island and deployment depth, see [@tbl:species_table_s].

![Representative photographs of the four most abundant crab species found across all deployed ARMS units (specimens shown were not collected as part of this study). (a) *Chlorodiella laevissima* (FLMNH 45490, Kāne&#x02BB;ohe, Hawai&#x02BB;i). (b) *Perinia tumida* (FLMNH 45652, Kāne&#x02BB;ohe, Hawai&#x02BB;i). (c) *Percnon abbreviatum* (FLMNH 45628, Kāne&#x02BB;ohe, Hawai&#x02BB;i). (d) *Dynomene hispida* (FLMNH 69204, Manawai, Hawai&#x02BB;i). Photo credits: Florida Museum of Natural History, licensed [CC BY-NC 3.0](https://creativecommons.org/licenses/by-nc/3.0/)](data/images/4crabs.png){#fig:fig_crabs}

<br><br>

### Mesophotic ARMS {-}

The `r nsamples(cc$crabs_deep_untransformed)` units deployed at mesophotic depths $\geq$ 30 m on O&#x02BB;ahu yielded fewer species and individuals compared to the complete dataset, with `r num(sum(sample_sums(cc$crabs_deep_untransformed)),1)` crabs overall, `r sum(unlist(deep_summary$types_breakdown))` distinct taxa (`r deep_summary$types_breakdown$species` species-level identifications, `r deep_summary$types_breakdown$genus` genus, `r deep_summary$types_breakdown$family` family), `r num(deep_summary$abundance_mean,0.1)` $\pm$ `r num(deep_summary$abundance_sd,0.1)` individuals per unit, and abundance ranging from `r format_range(deep_summary$abundance_range)`. The four most abundant taxa in the deep ARMS did not overlap with the most abundant taxa from the complete dataset: `r with(deep_summary,top_list(top,top_n,top_units))`. There were `r deep_summary$singleton_count` singletons, which proportionally accounted for ~`r pc(deep_summary$singleton_count/ntaxa(cc$crabs_deep_untransformed),0.1)` of species found on deep ARMS.

### Comparisons {-}
```{r depth-ranges, results="hide", dependson="taxonomy-validation"}
f <- cc$crabs %>%
  psmelt() %>%
  group_by(depth_zone) %>%
  filter(Abundance > 0) %>%
  ungroup() %>%
  group_by(family) %>%
  filter(n_distinct(shallow_deep) == 1) %>%
  ungroup()

fam <- f %>%
  distinct(family,shallow_deep) %>%
  count(shallow_deep) %>%
  deframe()

deep_fams <- f %>%
  filter(shallow_deep == "deep") %>%
  pull(family) %>%
  unique() %>%
  sort()

g <- cc$crabs %>%
  psmelt() %>%
  filter(!is.na(genus)) %>%
  group_by(depth_zone) %>%
  filter(Abundance > 0) %>%
  ungroup() %>%
  group_by(genus) %>%
  filter(n_distinct(shallow_deep) == 1) %>%
  ungroup()

gen <- g %>%
  distinct(genus,shallow_deep) %>%
  count(shallow_deep) %>%
  deframe()

deep_gens <- g %>%
  filter(shallow_deep == "deep") %>%
  pull(genus) %>%
  unique() %>%
  sort()

s <- cc$crabs %>%
  psmelt() %>%
  group_by(depth_zone) %>%
  filter(Abundance > 0) %>%
  ungroup() %>%
  group_by(species) %>%
  filter(n_distinct(shallow_deep) == 1) %>%
  ungroup()

spp <- s %>%
  distinct(species,shallow_deep) %>%
  count(shallow_deep) %>%
  deframe()

deep_spp <- s %>%
  filter(shallow_deep == "deep") %>%
  pull(species) %>%
  unique() %>%
  sort()
```
Out of the `r all_summary$taxonomy_breakdown$families` families recovered, `r fam['shallow']` were exclusive to shallow ARMS recovered from `r dr(cc$crabs_shallow_untransformed)` m and `r fam['deep']` were exclusive to units $\geq$ 30 m (`r format_list(deep_fams)`; [@tbl:family_depth_table_s]). `r str_to_sentence(english(gen['shallow']))` genera were found exclusively shallow and `r gen['deep']` exclusively deep (`r format_list(deep_gens)`; [@tbl:genus_depth_table_s]). `r str_to_sentence(english(spp['shallow']))` distinct taxa (including those collapsed to genus or family) were exclusive to shallow ARMS units and `r spp['deep']` were found only deep (`r format_list(deep_spp)`; [@tbl:species_depth_table_s]).

## Alpha diversity {-}

### Summary {-}

Overall mean species richness for the complete dataset was `r num(all_alpha$richness,0.1)` $\pm$ `r num(all_alpha$richness_sd,0.1)` species per unit (effective Simpson diversity: `r num(all_alpha$simpson,0.01)` $\pm$ `r num(all_alpha$simpson_sd,0.01)`). Overall mean species richness for shallow ARMS \< 30 m was `r num(shallow_alpha$richness,0.1)` $\pm$ `r num(shallow_alpha$richness_sd,0.1)` species per unit (effective Simpson diversity: `r num(shallow_alpha$simpson,0.01)` $\pm$ `r num(shallow_alpha$simpson_sd,0.01)`). Overall mean species richness for mesophotic ARMS $\geq$ 30 m was `r num(deep_alpha$richness,0.1)` $\pm$ `r num(deep_alpha$richness_sd,0.1)` species per unit (effective Simpson diversity: `r num(deep_alpha$simpson,0.01)` $\pm$ `r num(deep_alpha$simpson_sd,0.01)`).

### Comparisons {-}

For the complete dataset, there was no significant difference in observed species richness (`r k_test(all_alpha$k_r_island)`; [@fig:fig_alpha[]{}a]) or effective Simpson diversity (`r k_test(all_alpha$k_s_island,plain=TRUE)`; [@fig:fig_alpha[]{}c]) between the MHI and NWHI. Richness was not significantly different between shallow and deep samples (`r k_test(all_alpha$k_r_depth,plain=TRUE)`; [@fig:fig_alpha[]{}b]) but effective Simpson diversity was (`r k_test(all_alpha$k_s_depth,plain=TRUE)`; [@fig:fig_alpha[]{}d]). When comparing diversity across sampling depths for only sites in the MHI, richness was not significantly different (`r k_test(all_alpha$k_r_depth_main,plain=TRUE)`; [@fig:fig_alpha_main[]{}a]) but effective Simpson diversity was (`r k_test(all_alpha$k_s_depth_main,plain=TRUE)`; [@fig:fig_alpha_main[]{}c]). Neither richness nor Simpson diversity were significantly different between shallow and deep when comparisons were constrained to the island of O&#x02BB;ahu (Richness: `r k_test(all_alpha$k_r_depth_oahu,plain=TRUE)`, [@fig:fig_alpha_main[]{}b]; Simpson: `r k_test(all_alpha$k_s_depth_oahu,plain=TRUE)`, [@fig:fig_alpha_main[]{}d]). When only shallow arms \< 30 m were considered, there were no significant differences in alpha diversity between the MHI and NWHI for either index (richness: `r k_test(shallow_alpha$k_r_island,plain=TRUE)`, [@fig:fig_alpha2[]{}a]; effective Simpson: `r k_test(shallow_alpha$k_s_island,plain=TRUE)`, [@fig:fig_alpha2[]{}b]).

<!-- ### Effects of environmental variables {-} -->
## Brachyuran community structure {-}

```{r dbrda-analysis, results="hide", dependson="init"}
library(formula.tools)

# set random seed for consistency
set.seed(32767)

reduce_model <- function(m,o,d,vif_limit=3,condition=character()) {
  vifs <- vif.cca(m)
  if (any(vifs > vif_limit)) {
    oldterms <- rhs.vars(m$call$formula)
    v <- names(vifs)[which.max(vifs)]
    newterms <- grep(v,oldterms,value=TRUE,invert=TRUE)
    f <- reformulate(newterms,response=quote(o))
    newmod <- dbrda(f,data=d,distance="bray")
    reduce_model(newmod,o,d,vif_limit,condition)
  } else {
    return(m)
  }
}

# calculate full dbRDA model
db_modl <- dbrda(
  cc$crab_otus_shallow ~ depth +
    chl + sst + slope + coral_cover +
    larval_immigration + human_impact +
    lat + lon +
    Condition(recovery_year),
  data=cc$crab_data_shallow,
  # data=dd,
  distance="bray"
)

# reduce model to terms with VIF <= 3
db_modl <- reduce_model(db_modl,cc$crab_otus_shallow,cc$crab_data_shallow,vif_limit = 3)

# calculate biplot vectors and magnitudes
vectors <- db_modl %>%
  scores(display="bp") %>%
  as_tibble(rownames="term") %>%
  inner_join(model_terms %>% select(variable,driver=`Environmental Driver`),by=c("term" = "variable")) %>%
  mutate(size = sqrt(dbRDA1^2 + dbRDA2^2), driver=str_to_lower(driver)) %>%
  arrange(desc(size))

# get marginal term information and sort by descending f-statistic
modl_perm <- anova(db_modl,by="margin",permutations = PERM) %>%
  # as_tibble(rownames="term") %>%
  # janitor::clean_names() %>%
  broom::tidy() %>%
  arrange(desc(statistic))

# make a nicer display-worthy table for significant terms
rda_table <- modl_perm %>%
  inner_join(model_terms %>% select(variable,driver=`Environmental Driver`),by=c("term" = "variable")) %>%
  select(-term) %>%
  select(Term=driver,DF=df,`Sum of Squares`=SumOfSqs,`F-statistic`=statistic,`p-value`=p.value)

# get number of model terms
tt <- attr(db_modl$terms,"term.label")
nterm <- length(grep("Condition",tt,ignore.case = FALSE,value=TRUE,invert = TRUE))

# get top drivers for the text
drivers <- rda_table %>%
  slice(1:2) %>%
  select(term=Term,stat=`F-statistic`,pval=`p-value`)  %>%
  mutate(
    term=str_to_lower(term),
    term = str_replace_all(term,r'(\b[a-z]\b)',replacement=str_to_upper)
  )

all_terms <- english::english(nrow(model_terms))
```
### Environmental drivers {-}
Environmental variables were strongly clustered according to island group (MHI vs NWHI) ([@fig:fig_pca]). The best dbRDA model (`r permanova(db_modl,stat="F")`) explained `r pc(RsquareAdj(db_modl)$adj.r.squared,0.01)` (adjusted $\mathrm{R}^2$) of the variability in crab communities across the archipelago ([@fig:fig_dbrda]). `r str_to_sentence(english(nterm))` of the `r str_to_lower(english(nrow(model_terms)))` environmental variables tested were significant and sufficiently uncorrelated. The two drivers contributing most to explained variation were `r drivers$term[1]` (F = `r num(drivers$stat[1],0.01)`) and `r drivers$term[2]` (F = `r num(drivers$stat[2],0.01)`) ([@tbl:rda_table_s]). These factors also produced the longest effect vectors in the dbRDA plot ([@fig:fig_dbrda]).

```{r fig-dbrda, fig.cap="dbRDA plot showing significant drivers of crab community variation across the Hawaiian Islands. Colored ellipses represent 95% confidence intervals around island groups.", results="asis", fig.width=9, fig.height=6, dependson="rda-analysis"}
termz <- model_terms %>%
  select(variable,display=`Environmental Driver`)

# make the initial ordination plot
rda_plotz <- plot_ord(cc$crabs_shallow,db_modl,scale=2,color="island_group",labelsize=3,term_map=termz) + 
  stat_ellipse(geom="polygon",aes(group=island_group,fill=island_group),color=NA,alpha=0.1,show.legend = FALSE) +
  geom_point(aes(fill=island_group),shape=21,color="black",size=3) +
  plotz_theme() + 
  scale_fill_manual(name="Island group",values=island_group_pal,labels = c("Main","Northwest")) +
  guides(color = "none")

# calculate variation explained (constrained & unconstrained)
eigs <- eigenvals(db_modl)
constrained <- eigs[str_detect(names(eigs),"dbRDA")]
constrained <- constrained / sum(constrained)
unconstrained <- eigs/sum(eigs)

# function to format axis titles
splained <- function(name,constrained,total,accuracy=0.1) {
  str_glue("{name} ({pc(constrained,accuracy)} constrained, {pc(total,accuracy)} total)")
}

axes <- str_glue_data(list(num=1:2),"dbRDA{num}")
axis_var <- splained(axes,constrained[1:2],unconstrained[1:2])

# put the arrows on top
rda_plotz$layers <- rda_plotz$layers[c(1,2,5,6,3,4)]

# plot the plotz
rda_plotz + 
  xlab(axis_var[1]) + ylab(axis_var[2])
cat("{#fig:fig_dbrda}")
```

<br><br>

### Regional & depth comparisons {-}

```{r permanova, dependson="init"}
anova_region <- adonis2(cc$crab_dist_shallow ~ island_group,data=cc$crab_data_shallow,permutations=PERM)
anova_island <- adonis2(cc$crab_dist_shallow ~ island,data=cc$crab_data_shallow,permutations=PERM)
bd_region <- betadisper(cc$crab_dist_shallow,cc$crab_data_shallow$island_group)

anova_depth <- adonis2(cc$crab_dist ~ depth_zone,data=cc$crab_data,permutations=PERM)
anova_depth_oahu <- adonis2(cc$crab_dist_oahu ~ depth_zone,data=sample_tibble(cc$crabs_oahu),permutations=PERM)
anova_depth2_oahu <- adonis2(cc$crab_dist_oahu ~ shallow_deep,data=sample_tibble(cc$crabs_oahu),permutations=PERM)
bd_depth <- betadisper(cc$crab_dist,cc$crab_data$depth_zone)

anova_both <- adonis2(cc$crab_dist ~ island_group + depth_zone, data=cc$crab_data,permutations = PERM, by="margin")
effects <- broom::tidy(anova_both) %>%
  filter(!term %in% c("Residual","Total")) %>%
  pull(R2)
effect_ratio <- max(effects)/min(effects)

cd <- cc$crabs_deep %>%
  sample_tibble()
bd_deep <- betadisper(distance(cc$crabs_deep,"bray"),cd$depth_zone)
```

```{r cap-analysis,dependson="init",echo=FALSE,message=FALSE,warning=FALSE,results="hide"}
# convert otu tables and sample data into appropriate format
otus <- otu_table(cc$crabs) %>%
  as.data.frame()
cd <- cc$crab_data %>%
  mutate(sample_group = factor(sample_group,ordered=FALSE)) %>%
  as.data.frame()

# run CAP analysis, taking care of negative eigenvalues where necessary
cap_depth <- CAPdiscrim(otus ~ sample_group, data=cd, dist="bray", axes=2, m=0, add=TRUE)
# # calculate proportion of explained variance
cap_var <- cap_depth$manova$Eigenvalues / sum(cap_depth$manova$Eigenvalues)
cap_var <- list(x = cap_var[1], y = cap_var[2])

# make the sample group look nice
nice <- function(s) {
  str_replace(s,"shallow","Shallow") %>%
    str_replace("main","(MHI)") %>%
    str_replace("nwhi","(NWHI)") %>%
    str_replace("(^[0-9]+)","O‘ahu \\1") %>%
    str_split("((?<=[0-9])(?=[a-zA-Z]))|(_)") %>%
    map_chr(~str_c(.x,collapse=" "))
}

# get all the stuff we need to plot
pcap <- cap_depth$x %>%
  as_tibble(rownames="sample") %>%
  left_join(cd,by="sample") %>%
  mutate(
    sample_group = nice(sample_group),
    sample_group = fct_reorder(sample_group,depth)
  )

shallow_cd <- pcap %>%
  filter(shallow_deep == "shallow") %>%
  select(sample,LD1,LD2) %>%
  column_to_rownames("sample") %>%
  centroid_dist(mean=FALSE)

deep_cd <- pcap %>%
  filter(shallow_deep == "deep") %>%
  select(sample,LD1,LD2) %>%
  column_to_rownames("sample") %>%
  centroid_dist(mean=FALSE)

centroid_t <- t.test(deep_cd,shallow_cd,alternative = "g",var.equal = FALSE)

oahu_dist <- pcap %>%
  filter(island == "Oahu") %>%
  select(sample,LD1,LD2) %>%
  column_to_rownames("sample") %>%
  dist() %>% 
  c()

shallow_dist <- pcap %>%
  filter(shallow_deep == "shallow") %>%
  select(sample,LD1,LD2) %>%
  column_to_rownames("sample") %>%
  dist() %>% 
  c()

deep_dist <- pcap %>%
  filter(shallow_deep == "deep") %>%
  select(sample,LD1,LD2) %>%
  column_to_rownames("sample") %>%
  dist() %>%
  c()

distance_t <- t.test(deep_dist,shallow_dist,alternative = "g",var.equal = FALSE)
oahu_t <- t.test(oahu_dist,shallow_dist,alternative = "g",var.equal = FALSE)
```
Shallow crab community structure was significantly different between island groups (MHI vs. NWHI) (`r permanova(anova_region,0.01)`) as well as among individual islands (`r permanova(anova_island,0.01)`; see [@fig:fig_pairwise] for results of pairwise comparisons). There was no significant difference in community dispersion between the MHI and NWHI (`r permdisp(bd_region)`). When mesophotic ($\geq$ 30 m) ARMS units were included, there were significant differences among depth zones (\< 30 m, 30 m, 60 m, 90 m) both for all samples (`r permanova(anova_depth,0.01)`) and samples restricted to O&#x02BB;ahu (`r permanova(anova_depth_oahu,0.01)`), and community dispersion (`r permdisp(bd_depth)`). However, community dispersion was not different among depth zones $\geq$ 30 m (`r permdisp(bd_deep)`). Although crab communities were significantly different among both depth zones and island groups for the complete dataset (`r permanova(anova_both,0.01)`), depth explained `r num(effect_ratio,0.1)` times more variability than island group (based on model $\mathrm{R}^2$).

In the CAP analysis of island group and depth zone, ARMS units were correctly reassigned to sample groups `r pc(cap_depth$percent/100,0.1)` of the time. The first two linear discriminant axes explained `r pc(cap_var$x,0.1)` and `r pc(cap_var$y,0.1)` of the variation, respectively. Mesophotic samples (30--90 m) from O&#x02BB;ahu grouped separately from the shallow samples collected across the entire ~2,500 km expanse of the archipelago ([@fig:fig_cap]), whereas shallow samples from O&#x02BB;ahu primarily clustered with shallow samples from other islands ([@fig:fig_cap]). Variability among 30--90 m samples was apparently greater than among all shallow samples, with mean Euclidean distances between deep samples in the CAP ordination significantly greater than between shallow samples (`r t_test(distance_t)`; [@fig:fig_cap; @fig:fig_euclid]). Similarly, variability among samples from O&#x02BB;ahu (including mesophotic depths) exceeded variability among all shallow samples combined (based on Euclidean distance among CAP points, as above; `r t_test(oahu_t,plain=TRUE)`).

```{r fig-cap, fig.cap="CAP ordination of crab communities by island group and depth zone.", results="asis", fig.width=9, fig.height=6, dependson="cap-analysis"}
centroids <- pcap %>% group_by(sample_group) %>%
  summarise(LD1=mean(LD1),LD2=mean(LD2))
pcap <- pcap %>%
  mutate(
    oahu = factor(
      # if_else(island == "Oahu","oahu","noahu"),
      case_when(
        island == "Oahu" ~ "oahu",
        island == "French Frigate Shoals" ~ "ffs",
        .default = "noahu"
      ),
      levels = c("oahu","ffs","noahu")
    )
  )

ggplot(pcap,aes(x=LD1,y=LD2,fill=sample_group)) + 
  stat_ellipse(geom="polygon",level=0.95,alpha=0.1,show.legend = FALSE) +
  # geom_point(aes(shape=oahu),color="black",shape=21,size=5)  +
  geom_point(mapping=aes(shape=oahu),color="black",size=5)  +
  geom_label_repel(data=centroids,aes(label=sample_group), color="white", show.legend = FALSE) + 
  scale_shape_manual(values=c("oahu" = 23, "ffs" = 22, "noahu" = 21),name="Island", labels=c("noahu"="Other islands","oahu"="O‘ahu","ffs"="Lalo (French Frigate Shoals)" )) +
  scale_fill_manual(values=sample_depth_pal,name="Sample group") +
  guides( 
    fill=guide_legend(override.aes=list(shape=21)), 
    shape=guide_legend(override.aes=list(fill="black"))
  ) +
  xlab(str_glue("Linear discriminant 1 [{pc(cap_var$x,0.1)}]")) +
  ylab(str_glue("Linear discriminant 2 [{pc(cap_var$y,0.1)}]")) +
  plotz_theme()

# centroids <- pcap %>% group_by(sample_group) %>%
#   summarise(LD1=mean(LD1),LD2=mean(LD2))
# pcap <- pcap %>%
#   mutate(
#     oahu = factor(
#       if_else(island == "Oahu","oahu","noahu"),
#       levels = c("oahu","noahu")
#     )
#   )

# ggplot(pcap,aes(x=LD1,y=LD2,fill=sample_group)) + 
#   stat_ellipse(geom="polygon",level=0.95,alpha=0.1,show.legend = FALSE) +
#   # geom_point(aes(shape=oahu),color="black",shape=21,size=5)  +
#   geom_point(mapping=aes(shape=oahu),color="black",size=5)  +
#   geom_label_repel(data=centroids,aes(label=sample_group), color="white", show.legend = FALSE) + 
#   scale_shape_manual(values=c("oahu" = 23, "noahu" = 21),name="Island", labels=c("noahu"="Other islands","oahu"="O‘ahu")) +
#   scale_fill_manual(values=sample_depth_pal,name="Sample group") +
#   guides( 
#     fill=guide_legend(override.aes=list(shape=21)), 
#     shape=guide_legend(override.aes=list(fill="black"))
#   ) +
#   xlab(str_glue("Linear discriminant 1 [{pc(cap_var$x,0.1)}]")) +
#   ylab(str_glue("Linear discriminant 2 [{pc(cap_var$y,0.1)}]")) +
#   plotz_theme()
cat("{#fig:fig_cap}")
```

<br><br>

# Discussion {-}
  The Hawaiian Archipelago is among the most remote oceanic island chains on the planet and encompasses a range of gradients (anthropogenic, biotic, abiotic), making it an ideal natural laboratory to study the impacts of various drivers of species assemblages across space. Human activities are considerably higher in the populated MHI than in the protected and remote NWHI [@selkoeEvaluatingAnthropogenicThreats2008; @selkoeMapHumanImpacts2009; @selkoeDNACoralReef2016]. Although we did not consider the effects of individual impacts here, an index representing *cumulative* anthropogenic impacts was among the top drivers of shallow brachyuran community structure, similar to findings in previous work examining genetic diversity as it relates to species richness and habitat area, quality, and stability across the archipelago [@selkoeDNACoralReef2016]. The next-most significant driver, chlorophyll-A, can be used as a proxy for primary productivity. Productivity gradients across the Hawaiian Archipelago coincide with differences in coral reef community composition [@selkoeDNACoralReef2016], endemic reef fish dominance [@friedlanderDominanceEndemicsReef2020], survival rates of top predators [@bakerEffectVariableOceanic2007], and island geomorphology [@goveNearislandBiologicalHotspots2016]. Although not a significant factor in our model, coral cover and species composition are thought to affect patterns of benthic assemblages. Due to the limited availability of datasets covering our entire study area, our model only included coral cover at the island scale, but previous work has shown that coral species composition does not vary considerably across the archipelago [@griggCommunityStructureSuccession1983]. Crab communities from Lalo (French Frigate Shoals), which was found to harbor the highest coral species diversity [@griggCommunityStructureSuccession1983], did not group separately from the rest of our samples ([@fig:fig_cap]). Given the prevalence in our model of significant factors known to be important influencers of marine communities in Hawai&#x02BB;i, it is not surprising that we find the same top drivers of community structure for brachyuran crabs, one of the primary taxonomic groups inhabiting coral reefs. What is surprising---given the maximum vertical separation among shallow ARMS units of `r diff(range(cc$crab_data_shallow_unscaled$depth))` m---is the consistent significance of depth as a driver of community composition across these gradients.

Even though our shallow-water ARMS units were deployed within a narrow depth range (`r dr(cc$crabs_shallow_untransformed)` m), depth still emerged as a significant and consistent driver of community variation among these units. Depth in and of itself is unlikely to affect marine community composition. At the same time, oceanographic factors with well-known depth profiles (e.g., temperature, oxygen, nutrients, etc.) are not expected to vary dramatically across such a narrow depth range (\< 10 m), although fine-scale habitat variability among sites may also influence the community differences we found [@steyaertRemoteReefCryptobenthic2022]. Unfortunately, we lack environmental data at depth for individual deployments, and thus cannot unravel the co-varying factor(s) associated with the depth effects we observed. Regardless, given its significance across both narrow (\< 10 m) and wider (90 m) depth ranges sampled here, a factor associated with depth is clearly an important driver for cryptobenthic crab communities across the Hawaiian Archipelago. As brachyurans are a ubiquitous and important component of coral reef communities [@reynsPatternsProcessesBrachyuran1999; @lasleyjrPhylogeneticRelationshipsUbiquitous2015; @hazeriLatitudinalSpeciesDiversity2019], these findings are likely to apply more broadly as well.

While depth was a significant driver of community composition among shallow-water ARMS units ([@fig:fig_dbrda; @tbl:rda_table_s]), its influence was even more striking when comparing shallow to mesophotic samples. Due to logistical constraints (and the limitations of satellite-derived data products), our environmental data only covers shallow-water deployments, but CAP ordinations show considerable differences between the grouping of all shallow communities and the units deployed $\geq$ 30 m at the single island of O&#x02BB;ahu ([@fig:fig_cap]). According to these analyses, there was more community variation among samples spanning 90 m of depth on a single island than across 2,500 km of shallow habitat experiencing multiple gradients of environmental, ecological, and anthropogenic influence. The ordination revealed two distinct clusters for mesophotic samples: one comprising the 30 m samples and the other the 60 m and 90 m samples. These two clusters coincide with what is regarded as the upper and lower mesophotic zones, respectively [@kahngCommunityEcologyMesophotic2010; @rochaMesophoticCoralEcosystems2018; but see @pyleMesophoticCoralEcosystems2019]. The upper mesophotic (~30--60 m) has often been considered an extension of shallow reefs while the lower mesophotic (~60--150 m) is primarily thought to constitute a unique community. This pattern has held true across various taxonomic groups [@lesserGlobalCommunityBreaks2019; @rochaMesophoticCoralEcosystems2018], but our data show more distinct separations among shallow, upper, and lower mesophotic depths (although 30 m appears to encompass a transition zone).

In addition to community composition, mesophotic samples also exhibited alpha diversity differences. Effective Simpson diversity was significantly higher in deep communities (for comparisons among all islands and solely within the MHI, but not exclusively at O&#x02BB;ahu) ([@fig:fig_alpha; @fig:fig_alpha_main]). This pattern is in contrast to observations that biodiversity often decreases with depth [@pyleComprehensiveInvestigationMesophotic2016; @kahngCommunityEcologyMesophotic2010], although the differences we found were slight and our findings may be influenced by differences in habitat availability across depth zones (especially of the habitat types ARMS are designed to mimic).

Our depth transect was limited to O&#x02BB;ahu, but community and alpha diversity differences among depth zones at that single island appear to overwhelm differences observed among shallow-water coral reef habitat across the archipelago. Similar patterns were found at a local scale (tens of kilometers) on Hawai&#x02BB;i Island using environmental DNA (eDNA) [@hobanPlumbingDepthsEnvironmental2023]. In that study, communities were found to have more affinity within depth zones across sites than among depth zones within sites. These results and the patterns we observed suggest that depth-associated factors exert stronger ecological influence than those that vary across the latitudinal, physical, and anthropogenic gradients throughout the Hawaiian Archipelago.

Since we lack environmental data for mesophotic samples, we cannot make conclusions about the specific influences driving the differences we found. However, light and temperature have been proposed as important bottom-up drivers of depth zonation in benthic communities [@kahngCommunityEcologyMesophotic2010; @lesserGlobalCommunityBreaks2019; @pyleComprehensiveInvestigationMesophotic2016; @pyleMesophoticCoralEcosystems2019]. Oceanographic conditions may be inconsistent close to shore, but the average mixed-layer depth calculated (using 0.8 &deg;C $\upDelta$T [@karaOptimalDefinitionOcean2000]) from CTD casts taken at station ALOHA (22.75&deg;N, 158.00&deg;W; within the latitudinal expanse of the MHI) was 84.6 $\pm$ 26 m over the course of the year (based on monthly averages from casts taken across the duration of ARMS unit deployments [@HawaiiOceanTimeseries]). Thus, given the assumption of similar temperature profiles near the O&#x02BB;ahu shoreline, temperature alone appears insufficient to explain the zonation patterns we observed---although due to its significant role as a driver of shallow community variation, it cannot be ruled out. In contrast, the lower depth boundaries of mesophotic ecosystems are largely defined by surface irradiance [@kahngCommunityEcologyMesophotic2010; @hindersteinThemeSectionMesophotic2010], so light seems a possible candidate as bottom-up abiotic factor influencing community structure across depths. Crab communities are unlikely to be directly affected by light variability. However, habitat-building species such as corals, algae, and sponges may be [@kahngCommunityEcologyMesophotic2010; @slatteryFunctionStabilityMesophotic2024a], and may in turn affect the structure of the communities dependent upon them [@hustonCompetitionCoexistenceEffects1994; @smithRoleHabitatComplexity2014; @sebensHabitatStructureCommunity1991]. If patterns of community structure are driven primarily from the bottom up by light, that may contribute to our observation of a significant depth effect over less than 10 m in our shallow samples.

Further work to investigate drivers of depth zonation could take advantage of established gradients and use the Hawaiian Archipelago as a natural laboratory, by sampling depth transects across the archipelago. Coral reef biodiversity is generally thought to be structured at depth by a combination of the attenuation of light, a decrease in temperature, increasing salinity, and a reduction in trophic resources [@lesserEcologyMesophoticCoral2009; @kahngCommunityEcologyMesophotic2010; @slatteryFunctionStabilityMesophotic2024a]. If mesophotic communities are found to be different from their shallow counterparts across multiple islands (ideally comparing sites experiencing similar irradiance), this suggests that they are vertically structured by a factor (e.g., light) outside of the numerous natural gradients that exist across the Hawaiian Islands. In contrast, other factors (such as temperature, dissolved inorganic nitrogen, particulate organic carbon, or productivity) expected to have more strongly-defined gradients across the regions [@slatteryFunctionStabilityMesophotic2024a], would lead to variable zonation patterns that should be detected in our spatial comparison of the MHI & NWHI. Such differences should be tested in future studies to understand the specific physical and ecological drivers of coral reef depth zonation. The logistical complexity and time investment required for ARMS deployments make this challenging, but emerging approaches such as environmental DNA (eDNA) may offer a suitable alternative.

Marine communities are structured by a suite of factors, but there remains no consensus for either generalized mechanisms or the relative importance of drivers of community structure for coral reefs. Our best knowledge of the role of depth in structuring marine communities comes from the intertidal, and points to biotic factors (e.g., predation, competition) determining lower limits and abiotic factors (e.g., desiccation, thermal stress) setting upper limits of vertical species distributions [@connellEffectsCompetitionPredation1961; @connellInfluenceInterspecificCompetition1961; @dennyPhysicalProcessesThat2001; @paineIntertidalCommunityStructure1974]. There are numerous examples from other study systems in which the relative contributions of biotic versus abiotic factors to structuring communities run in opposing directions across some habitat gradient [e.g., climate, soil type, elevation, etc. @bradyEvolutionaryEcologyPlant2005; @moore10PlantCompetition2011; @wishingradTemperateZoneIsolation2023]. However, abiotic stressors on intertidal communities are obviously different from those on coral reefs, and generalizations about the mechanisms and ecological processes that structure coral reef communities remain a subject of considerable debate [@bellwoodEnvironmentalGeometricConstraints2005; @bellwoodSleepingFunctionalGroup2006; @carrBiodiversityPopulationRegulation2002; @connollyIndoPacificBiodiversityCoral2003; @dornelasCoralReefDiversity2006; @grahamImportanceStructuralComplexity2013; @komyakovaRelativeImportanceCoral2013]. In the Hawaiian archipelago, we found significant differences in the composition of shallow coral reef crab communities between the populated high islands of the MHI and the uninhabited atolls of the NWHI. Moreover, these differences were driven by a combination of environmental (sea-surface temperature, chlorophyll-A concentration, depth), biotic (potential larval immigration), and anthropogenic (cumulative human impacts) factors. Despite the significance of these factors in driving shallow community variation across the archipelago, mesophotic assemblages were found to be distinct ([@fig:fig_cap]). Although there were variables we did not consider in our model, this suggests that vertical distributions are influenced by factors other than those driving spatial variation across the island chain. In light of this, and since mesophotic communities are generally less biodiverse than shallow reefs [@kahngCommunityEcologyMesophotic2010; @lesserEcologyMesophoticCoral2009; @pyleComprehensiveInvestigationMesophotic2016], it may be possible that the gradients driving depth distributions on coral reefs run opposite to those in the intertidal. Given the rapid changes in important factors like temperature, nutrients, and light throughout the upper ~100--150 m of the water column (particularly the diminishment of photosynthetically active radiation below 100 m), the *lower* limits of the coral reef mesophotic zone may be determined by abiotic factors that correlate with depth. Conversely, the *upper* limits of the mesophotic zone (the transition from shallow to mesophotic habitat), which occur at the higher end of the depth-biodiversity gradient [@kahngCommunityEcologyMesophotic2010; @lesserEcologyMesophoticCoral2009] may be driven by ecological interactions [@vellendEffectsDiversityDiversity2008].

By comparing shallow-water ARMS across the latitudinal, physical and anthropogenic gradients of the Hawaiian Archipelago, we were able to test the relative importance of the significant drivers of community structure. We additionally find that a 90 m depth transect at a single island shows greater divergence in canonical space than the full 2,500 km gradient of the Hawaiian Archipelago for shallow-water ARMS. Our analyses corroborate previous work and show that depth is among the strongest drivers of community structure. We conclude that some factor(s) covarying with depth exert more influence on cryptobenthic coral reef community structure than those biotic and abiotic gradients that span the entire Hawaiian Archipelago in shallow-water environments.

# Acknowledgements
We thank the NOAA Coral Reef Conservation Program for supporting the deployment and recovery of the shallow ARMS. We thank the UH Mānoa Dive Safety Office (D. Pence, K. Stender, J. Jones, and C.J. Bradley) for installation and retrieval of the mesophotic ARMS. We specially thank K. Flanagan from the UH dive program, whose support was instrumental in the development of the mesophotic ARMS project, and J. Copus, who provided expert mesophotic field support; both were taken too soon. May their memories be a blessing. We thank the members of the ToBo lab, particularly M. Iacchei, C. Ka&#x02BB;apu-Lyons, and Z. Hee, for discussion, support, and their efforts in the disassembly and sorting of ARMS samples on retrieval. Finally, we are grateful to two anonymous reviewers for constructive comments and suggestions on an earlier version of this paper. This is contribution #xxx from the Hawai&#x02BB;i Institute of Marine Biology and #xxx from the School of Ocean and Earth Science and Technology at the University of Hawai&#x02BB;i.

# Funding
This work was funded through a combination of grants including NSF OCE-1260169, NSF GRFP DGE-1329626, and NSF OCE-2048457. In addition, the NOAA Coral Reef Conservation Program funded the Pacific Reef Assessment and Monitoring Program, which included the shallow ARMS deployment in the Hawaiian Archipelago.
<!--
Specimens used in this study were collected under Hawai&#x02BB;i Division of Aquatic Resources Special Activity Permit numbers SAP 2013-47, SAP 2012-63, SAP 2011-01, and SAP 2008-99.
-->

# References {-}

<!-- show main bibliography -->
<div class="multi-refs"></div>

\newpage

# Supplemental material {label="S"}

## Data import and processing

All data import and processing code for environmental variables used in the dbRDA model can be found in the file `get-data.R` from the associated github repository. Details for specific datasets are given below.

### Processing of satellite-derived environmental data
Level-3 mapped (4 km grid-size) sea-surface temperature (SST) and chlorophyll-A (chl-A) data were retrieved from NASA's Physical Oceanography Distributed Active Archive Center [PO.DAAC; @nasaobpgMODISAquaGlobal2020] and Ocean Biology Distributed Active Archive Center [OB.DAAC; @nasagoddardspaceflightcenterModerateresolutionImagingSpectroradiometer2022] respectively. Monthly summary datasets were downloaded for the period of 2009--2016 and averaged over each of the ARMS unit soaking periods (e.g. 2009--2011, 2010--2012, etc.). Values (SST or chl-A) were extracted for the geographic coordinates of each sampling site from the averaged dataset for the relevant soaking period. Original raw monthly-average dataset names can be found in @tbl:nasa_table_s.

### Processing of cumulative human impacts data
Gridded cumulative human impacts raster data calculated by Halpern et al [-@halpernRecentPaceChange2019] was downloaded from the Knowledge Network for Biocomplexity (KNB) repository [@frazierRecentPaceChange2019]. Human impacts datasets were downloaded for each available year in this study (2009--2013) and averaged over ARMS unit soaking periods (as above). Since cumulative impacts data was only available through 2013, we used the combined average of all available years for the 2013--2016 soaking period. Since the availability of data points at individual coordinates was highly variable, we averaged impact scores for each sampled island in the study.

### Other datasets
Coral cover and potential larval immigration were imported from the online supplemental material accompanying Selkoe et al. [-@selkoeDNACoralReef2016].

\newpage

## Figures and tables

```{r fig-alpha, fig.cap='Box plots of (a,b) species richness and (c,d) effective Simpson diversity for complete dataset (all sampled depths). Asterisks indicate significant differences (K-W test). Notches represent nonparametric confidence intervals about the median. (a) Species richness by island group. (b) Species richness by sampling depth. (c) Effective Simpson diversity by island group. (d) Effective Simpson diversity by sampling depth.', results="asis", dev="svg", fig.width=10, fig.height=10, dependson="init"}
library(patchwork)

art <- all_alpha$richness_table %>% 
  mutate(island_group = str_to_sentence(island_group)) %>%
  mutate(shallow_deep = str_to_sentence(shallow_deep),shallow_deep = fct_reorder(shallow_deep,depth))

if (all_alpha$k_r_island$p.value < 0.05) {
  which_max <- art %>% 
    group_by(island_group) %>% 
    summarise(r = mean(Observed)) %>%
    slice(which.max(r)) %>%
    pull(island_group) %>%
    as.character()
  max_r <- max(art$Observed) + (diff(range(art$Observed)))/10
  sig <- tribble(
    ~x,~y,
    which_max,max_r
  ) 
} else {
  sig <- tribble(~x,~y)
}

pointsize <- 5

p1 <- ggplot(art) + 
  geom_boxplot(aes(x=island_group,y=Observed,fill=island_group),notch=TRUE) +
  geom_point(aes(x,y),data=sig,shape=8,size=pointsize,stroke=1.5) + 
  scale_fill_manual(values=island_group_pal) +
  guides(fill="none") + 
  plotz_theme() + 
  xlab("Island group") + 
  ylab("Richness")

if (all_alpha$k_r_depth$p.value < 0.05) {
  which_max <- art %>% 
    group_by(shallow_deep) %>% 
    summarise(r = mean(Observed)) %>%
    slice(which.max(r)) %>%
    pull(shallow_deep) %>%
    as.character()
  max_r <- max(art$Observed) + (diff(range(art$Observed)))/10
  sig <- tribble(
    ~x,~y,
    which_max,max_r
  ) 
} else {
  sig <- tribble(~x,~y)
}

p2 <- ggplot(art) + 
  geom_boxplot(aes(x=shallow_deep,y=Observed,fill=shallow_deep),notch=TRUE) +
  geom_point(aes(x,y),data=sig,shape=8,size=pointsize,stroke=1.5) + 
  scale_fill_manual(values=sample_depth_pal) +
  guides(fill="none") + 
  plotz_theme()  + 
  xlab("Sampling depth") + 
  ylab("Richness")

if (all_alpha$k_s_island$p.value < 0.05) {
  which_max <- art %>% 
    group_by(island_group) %>% 
    summarise(r = mean(Simpson)) %>%
    slice(which.max(r)) %>%
    pull(island_group) %>%
    as.character()
  max_r <- max(art$Simpson) + (diff(range(art$Simpson)))/10
  sig <- tribble(
    ~x,~y,
    which_max,max_r
  ) 
} else {
  sig <- tribble(~x,~y)
}

pointsize <- 5

p3 <- ggplot(art) + 
  geom_boxplot(aes(x=island_group,y=Simpson,fill=island_group),notch=TRUE) +
  geom_point(aes(x,y),data=sig,shape=8,size=pointsize,stroke=1.5) + 
  scale_fill_manual(values=island_group_pal) +
  guides(fill="none") + 
  plotz_theme() + 
  xlab("Island group") + 
  ylab("Effective Simpson diversity")

if (all_alpha$k_s_depth$p.value < 0.05) {
  which_max <- art %>% 
    group_by(shallow_deep) %>% 
    summarise(r = mean(Simpson)) %>%
    slice(which.max(r)) %>%
    pull(shallow_deep) %>%
    as.character()
  max_r <- max(art$Simpson) + (diff(range(art$Simpson)))/10
  sig <- tribble(
    ~x,~y,
    which_max,max_r
  ) 
} else {
  sig <- tribble(~x,~y)
}

p4 <- ggplot(art) + 
  geom_boxplot(aes(x=shallow_deep,y=Simpson,fill=shallow_deep),notch=TRUE) +
  geom_point(aes(x,y),data=sig,shape=8,size=pointsize,stroke=1.5) + 
  scale_fill_manual(values=sample_depth_pal) +
  guides(fill="none") + 
  plotz_theme()  + 
  xlab("Sampling depth") + 
  ylab("Effective Simpson diversity")

(p1 + p2) / (p3 + p4) & 
  plot_annotation(tag_levels = list(str_glue("({letters[1:4]})"))) & 
  theme(plot.tag = element_text(face="italic"))
cat("{#fig:fig_alpha}")
```

\newpage

```{r fig-alpha-main, fig.cap='Box plots of species richness and effective Simpson diversity per depth zone for Main Hawaiian Islands (a,c) and O&#x02BB;ahu (b,d) only. Asterisks and notches as above. (a) Species richness by sampling depth, MHI. (b) Species richness by sampling depth, O&#x02BB;ahu. (c) Effective Simpson diversity by sampling depth, MHI. (d) Effective Simpson diversity by sampling depth, O&#x02BB;ahu.', results="asis", dev="svg", fig.width=10, fig.height=10, dependson="init"}
# depth comparison for main islands only
art <- all_alpha$richness_table %>% 
  filter(island_group == "main") %>%
  mutate(island_group = str_to_sentence(island_group)) %>%
  mutate(shallow_deep = str_to_sentence(shallow_deep),shallow_deep = fct_reorder(shallow_deep,depth))

if (all_alpha$k_r_depth$p.value < 0.05) {
  which_max <- art %>% 
    group_by(shallow_deep) %>% 
    summarise(r = mean(Observed)) %>%
    slice(which.max(r)) %>%
    pull(shallow_deep) %>%
    as.character()
  max_r <- max(art$Observed) + (diff(range(art$Observed)))/10
  sig <- tribble(
    ~x,~y,
    which_max,max_r
  ) 
} else {
  sig <- tribble(~x,~y)
}

pointsize <- 5

p1 <- ggplot(art) + 
  geom_boxplot(aes(x=shallow_deep,y=Observed,fill=shallow_deep),notch=TRUE) +
  geom_point(aes(x,y),data=sig,shape=8,size=pointsize,stroke=1.5) + 
  scale_fill_manual(values=sample_depth_pal) +
  guides(fill="none") + 
  plotz_theme()  + 
  xlab("Sampling depth - MHI") + 
  ylab("Richness")

if (all_alpha$k_s_depth$p.value < 0.05) {
  which_max <- art %>% 
    group_by(shallow_deep) %>% 
    summarise(r = mean(Simpson)) %>%
    slice(which.max(r)) %>%
    pull(shallow_deep) %>%
    as.character()
  max_r <- max(art$Simpson) + (diff(range(art$Simpson)))/10
  sig <- tribble(
    ~x,~y,
    which_max,max_r
  ) 
} else {
  sig <- tribble(~x,~y)
}

p2 <- ggplot(art) + 
  geom_boxplot(aes(x=shallow_deep,y=Simpson,fill=shallow_deep),notch=TRUE) +
  geom_point(aes(x,y),data=sig,shape=8,size=pointsize,stroke=1.5) + 
  scale_fill_manual(values=sample_depth_pal) +
  guides(fill="none") + 
  plotz_theme()  + 
  xlab("Sampling depth - MHI") + 
  ylab("Effective Simpson diversity")

art <- all_alpha$richness_table %>% 
  filter(island == "Oahu") %>%
  mutate(island_group = str_to_sentence(island_group)) %>%
  mutate(shallow_deep = str_to_sentence(shallow_deep),shallow_deep = fct_reorder(shallow_deep,depth))

if (all_alpha$k_r_depth_oahu$p.value < 0.05) {
  which_max <- art %>% 
    group_by(shallow_deep) %>% 
    summarise(r = mean(Observed)) %>%
    slice(which.max(r)) %>%
    pull(shallow_deep) %>%
    as.character()
  max_r <- max(art$Observed) + (diff(range(art$Observed)))/10
  sig <- tribble(
    ~x,~y,
    which_max,max_r
  ) 
} else {
  sig <- tribble(~x,~y)
}

pointsize <- 5

p3 <- ggplot(art) + 
  geom_boxplot(aes(x=shallow_deep,y=Observed,fill=shallow_deep),notch=TRUE) +
  geom_point(aes(x,y),data=sig,shape=8,size=pointsize,stroke=1.5) + 
  scale_fill_manual(values=sample_depth_pal) +
  guides(fill="none") + 
  plotz_theme()  + 
  xlab("Sampling depth - O‘ahu") + 
  ylab("Richness")

if (all_alpha$k_s_depth_oahu$p.value < 0.05) {
  which_max <- art %>% 
    group_by(shallow_deep) %>% 
    summarise(r = mean(Simpson)) %>%
    slice(which.max(r)) %>%
    pull(shallow_deep) %>%
    as.character()
  max_r <- max(art$Simpson) + (diff(range(art$Simpson)))/10
  sig <- tribble(
    ~x,~y,
    which_max,max_r
  ) 
} else {
  sig <- tribble(~x,~y)
}

p4 <- ggplot(art) + 
  geom_boxplot(aes(x=shallow_deep,y=Simpson,fill=shallow_deep),notch=TRUE) +
  geom_point(aes(x,y),data=sig,shape=8,size=pointsize,stroke=1.5) + 
  scale_fill_manual(values=sample_depth_pal) +
  guides(fill="none") + 
  plotz_theme()  + 
  xlab("Sampling depth - O‘ahu") + 
  ylab("Effective Simpson diversity")

(p1 + p3) / (p2 + p4) & 
  plot_annotation(tag_levels = list(str_glue("({letters[1:4]})"))) & 
  theme(plot.tag = element_text(face="italic"))
cat("{#fig:fig_alpha_main}")
```

\newpage

```{r fig-alpha2, fig.cap='Box plots of (a) species richness and (b) Effective Simpson diversity by island group for shallow samples only. Notches as in [@fig:fig_alpha]. Both comparisons were non-significant.', results="asis", dev="svg", fig.width=8, fig.height=4, dependson="init"}
art <- shallow_alpha$richness_table %>% 
  mutate(island_group = str_to_sentence(island_group)) %>%
  mutate(shallow_deep = str_to_sentence(shallow_deep),shallow_deep = fct_reorder(shallow_deep,depth))

if (shallow_alpha$k_r_island$p.value < 0.05) {
  which_max <- art %>% 
    group_by(island_group) %>% 
    summarise(r = mean(Observed)) %>%
    slice(which.max(r)) %>%
    pull(island_group) %>%
    as.character()
  max_r <- max(art$Observed) + (diff(range(art$Observed)))/10
  sig <- tribble(
    ~x,~y,
    which_max,max_r
  ) 
} else {
  sig <- tribble(~x,~y)
}

pointsize <- 5

p1 <- ggplot(art) + 
  geom_boxplot(aes(x=island_group,y=Observed,fill=island_group),notch=TRUE) +
  geom_point(aes(x,y),data=sig,shape=8,size=pointsize,stroke=1.5) + 
  scale_fill_manual(values=island_group_pal) +
  guides(fill="none") + 
  plotz_theme() + 
  xlab("Island group") + 
  ylab("Richness")

if (shallow_alpha$k_s_island$p.value < 0.05) {
  which_max <- art %>% 
    group_by(island_group) %>% 
    summarise(r = mean(Simpson)) %>%
    slice(which.max(r)) %>%
    pull(island_group) %>%
    as.character()
  max_r <- max(art$Simpson) + (diff(range(art$Simpson)))/10
  sig <- tribble(
    ~x,~y,
    which_max,max_r
  ) 
} else {
  sig <- tribble(~x,~y)
}

pointsize <- 5

p2 <- ggplot(art) + 
  geom_boxplot(aes(x=island_group,y=Simpson,fill=island_group),notch=TRUE) +
  geom_point(aes(x,y),data=sig,shape=8,size=pointsize,stroke=1.5) + 
  scale_fill_manual(values=island_group_pal) +
  guides(fill="none") + 
  plotz_theme() + 
  xlab("Island group") + 
  ylab("Effective Simpson diversity")

(p1 + p2) & 
  plot_annotation(tag_levels = list(str_glue("({letters[1:2]})"))) & 
  theme(plot.tag = element_text(face="italic"))
cat("{#fig:fig_alpha2}")
```

\newpage

```{r fig-pca, fig.cap='Principal component analysis (PCA) of environmental variables by sampling site. Ellipses represent 95% confidence intervals around island groups.', results='asis', dev='svg', fig.width=8,fig.height=8, dependson='init'}
cds <- cc$crab_data_shallow

pcdata <- cds %>%
  select(where(is.numeric),-c(lon,lat),-closest_island) %>%
  select(deployment_year,depth,chl,sst,slope:human_impact,-contains('radius'))

pca <- prcomp(pcdata)
explained <- (pca$sdev^2)/sum(pca$sdev^2)

bp <- cds %>%
  bind_cols(scores(pca))

loadings <- pca$rotation %>%
  as_tibble(rownames = "label") %>%
  mutate(
    PC1 = PC1*3,
    PC2 = PC2*3,
    hjust = if_else(PC1 < 0,1,0),
    xnudge = if_else(PC1 < 0,-0.02,0.02)
  )

ggplot(bp,aes(x=PC1,y=PC2)) + 
  geom_point(aes(fill=island_group),shape=21,color="black",size=4) +
  stat_ellipse(geom="polygon",aes(group=island_group,fill=island_group),color=NA,alpha=0.1,show.legend = FALSE) +
  # geom_mark_ellipse(color="black",alpha=0.2,show.legend=FALSE, linewidth=0.07) +
  scale_fill_manual(values=island_group_pal,name="Island group") + 
  geom_segment(data=loadings,aes(x=0,y=0,xend=PC1,yend=PC2),color="black",size=0.5,arrow=arrow(length=unit(0.02,"npc"))) + 
  geom_label(data=loadings,aes(x=PC1+xnudge,y=PC2,label=label,hjust=hjust)) + 
  xlab(str_glue("PC1 [{pc(explained[1])}]")) + 
  ylab(str_glue("PC2 [{pc(explained[2])}]")) + 
  plotz_theme()
cat("{#fig:fig_pca}")
```

\newpage

```{r pairwise-island, fig.cap='Results of pairwise PERMANOVA tests by island for shallow crab communities (p-values adjusted according to the false discovery rate [@benjaminiControllingFalseDiscovery1995]). Significant comparisons presented in bold.', results="asis", dev="svg", fig.width=12, fig.height=9, dependson="init"}
pairwise_island <- pairwise_adonis(cc$crab_dist_shallow,cc$crab_data_shallow$hawaiian_name,permutations=PERM)

pw <- pairwise_island %>%
  filter(p_adj < 0.05)

islands <- cc$crabs_shallow_untransformed %>%
  sample_tibble() %>%
  pull(hawaiian_name) %>%
  levels()

pw1 <- pairwise_island %>%
  select(left,right,p_adj,pseudo_f) %>%
  mutate(left=factor(left,levels=rev(islands)),right=factor(right,levels=rev(islands))) %>%
  mutate(sig = round(p_adj,2) < 0.05, p_adj = pval(p_adj), f = str_glue("F = {num(pseudo_f)}")) %>%
  mutate(
    combo = map2_chr(left,right,~str_c(sort(c(.x,.y)),collapse="-"))
  ) %>%
  select(-c(left,right)) %>%
  separate_wider_delim(combo,"-",names=c("left","right")) %>%
  select(left,right,everything()) %>%
  mutate(left=factor(left,levels=islands),right=factor(right,levels=islands)) %>%
  mutate(txt = str_glue("{p_adj}<br>{f}"))

ggplot(pw1) + 
  geom_richtext(
    aes(
      x=right,
      y=left,
      label=txt,
      fontface = if_else(sig,"bold","plain")
    ),
    fill = NA,
    label.color = NA
    # size=10
  ) +
  plotz_theme() +
  theme(
    panel.border = element_blank(),
    axis.line = element_line(color="black"),
    axis.text = element_text(size = 14),
    axis.text.x = element_text(angle=30,vjust=1,hjust=0)
  ) +
  guides(fill="none") +
  xlab("") + ylab("") +
  scale_x_discrete(position="top") +
  scale_y_discrete(limits=rev)
cat("{#fig:fig_pairwise}")
```

\newpage

```{r fig-euclid, fig.cap='Box plot of Euclidean distances between points in CAP ordination by depth zone. Asterisk indicates significant difference (Welch two-sample t-test).', results="asis", dev="svg", fig.width=6, fig.height=4, dependson="cap-analysis"}
dists <- bind_rows(
  tibble(dist=shallow_dist,depth="Shallow"),
  tibble(dist=deep_dist,depth="Deep")
) %>%
  mutate(depth = factor(depth,levels=c("Shallow","Deep")))

if (distance_t$p.value < 0.05) {
  which_max <- dists %>%
    group_by(depth) %>%
    summarise(d = mean(dist)) %>%
    slice(which.max(d)) %>%
    pull(depth) %>%
    as.character()
  max_d <- max(dists$dist) + (diff(range(dists$dist)))/10
  sig <- tribble(
    ~x,~y,
    which_max,max_d
  ) 
} else {
  sig <- tribble(~x,~y)
}

pointsize <- 5
ggplot(dists,aes(x=depth,y=dist,fill=depth)) + 
  geom_boxplot(notch=TRUE) + 
  geom_point(aes(x,y),data=sig,shape=8,size=pointsize,stroke=1.5,fill="black") +
  scale_fill_manual(values=sample_depth_pal) + 
  guides(fill="none") + 
  plotz_theme() + 
  xlab("Sampling depth") + 
  ylab("Euclidean distance between CAP points")
cat("{#fig:fig_euclid}")
```

\newpage

```{r arms-metadata-table, results="asis", dependson="init"}
region_map <- c(
  'main' = 'MHI',
  'northwest' = 'NWHI',
  'mce' = 'MHI (mesophotic)'
)

arms_metadata <- cc$crabs_untransformed %>%
  sample_tibble() %>%
  # separate_wider_regex(sample,patterns=c(recovery_year="^[0-9]+","_",sample=".*"))  %>%
  # select(sample,everything()) %>%
  mutate(
    island = case_when(
      str_remove_all(hawaiian_name,'[^a-zA-Z ]') != island ~ as.character(str_glue("{hawaiian_name} ({island})")),
      TRUE ~ as.character(hawaiian_name)
    ),
    island = case_when(
      island == "O‘ahu" ~ str_glue("{hawaiian_name} ({shallow_deep})"),
      TRUE ~ island
    ),
    island = fct_reorder(island,-lat),
    region = region_map[region],
    lat = num(lat,0.0001),
    lon = num(lon,0.0001)
  ) %>%
  select(sample,region,island,depth,deployment_year,recovery_year,lat,lon) %>%
  janitor::clean_names(case="sentence")
print_table(arms_metadata,"arms_metadata_table","ARMS unit metadata",accuracy=1,hdr_case=NA,format_numbers=FALSE) 
```

\newpage

```{r species-table, results="asis", dependson="taxonomy-validation"}
# make table of all sampled taxa
species_tbl <- cc$crabs %>%
  psmelt() %>%
  # sum counts by island group, depth zone, and species
  group_by(OTU,island_group,hawaiian_name,depth_zone,depth,family,genus,species) %>%
  summarise(n=sum(Abundance)) %>%
  ungroup() %>%
  # keep species we actually found
  filter(n > 0) %>%
  # figure out if a species is exclusive to a depth zone
  group_by(species) %>%
  mutate(zones = str_c(depth_zone,collapse="|")) %>%
  ungroup() %>%
  mutate(depth_exclusive = depth_zone == zones) %>%
  arrange(island_group,hawaiian_name,depth_zone,depth,family,genus,species) %>%
  # join in validated taxonomy
  left_join(crab_taxonomy_validated,by=c("OTU" = "otu")) %>%
  select(hawaiian_name,island_group,depth_zone,depth,family.y,display_species,depth_exclusive) %>%
  rename(island = hawaiian_name,valid_species=display_species) %>%
  mutate(
    across(c(island_group,depth_zone),str_to_sentence),
    depth_exclusive = if_else(depth_exclusive,"Yes","No")
  ) %>%
  rename_with(\(n) {
    n %>%
      str_replace_all("depth$","depth (m)") %>%
      str_replace_all("_"," ") %>%
      str_replace_all("\\.y$","") %>%
      str_replace_all("^display.","") %>%
      str_to_sentence()
  })
print_table(species_tbl,"species_table_s","List of crab taxa detected by island and ARMS unit deployment depth",hdr_case="none",accuracy=1)
```

\newpage

```{r family-depth-table, results="asis", dependson="taxonomy-validation"}
deep_map <- c(
  "shallow" = str_glue("Shallow ({dr(cc$crabs_shallow_untransformed)} m)"),
  "deep" = "Deep ($\\geq$ 30 m)"
)
crabbo <- cc$crabs %>%
  psmelt() %>%
  mutate(depth_zone = replace(depth_zone,which(depth_zone == "shallow"),str_c(dr(cc$crabs_shallow_untransformed)," m"))) %>%
  group_by(depth_zone) %>%
  filter(Abundance > 0) %>%
  ungroup() %>%
  group_by(family) %>%
  filter(n_distinct(shallow_deep) == 1) %>%
  summarise(zones = str_c(unique(depth_zone),collapse=", "), deeps = str_c(unique(shallow_deep),collapse=", "),species=str_c(unique(species),collapse=", ")) %>%
  ungroup() %>%
  arrange(deeps,zones,family,species) %>%
  select(deeps,family,species,zones) %>%
  mutate( deeps = deep_map[deeps] ) %>%
  rename(Family=family,`"Species"`=species,`Depth range`=deeps,`Depth zones`=zones)

print_table(crabbo,"family_depth_table_s","Brachyuran families exclusive to shallow or mesophotic habitat",hdr_case="none",accuracy=1)
```

\newpage

```{r genus-depth-table, results="asis", dependson="taxonomy-validation"}
deep_map <- c(
  "shallow" = str_glue("Shallow ({dr(cc$crabs_shallow_untransformed)} m)"),
  "deep" = "Deep ($\\geq$ 30 m)"
)
crabbo <- cc$crabs %>%
  psmelt() %>%
  mutate(depth_zone = replace(depth_zone,which(depth_zone == "shallow"),str_c(dr(cc$crabs_shallow_untransformed)," m"))) %>%
  group_by(depth_zone) %>%
  filter(Abundance > 0) %>%
  ungroup() %>%
  group_by(genus) %>%
  filter(n_distinct(shallow_deep) == 1) %>%
  summarise(zones = str_c(unique(depth_zone),collapse=", "), deeps = str_c(unique(shallow_deep),collapse=", "),species=str_c(unique(species),collapse=", ")) %>%
  ungroup() %>%
  arrange(deeps,zones,genus,species) %>%
  select(deeps,genus,species,zones) %>%
  mutate( deeps = deep_map[deeps] ) %>%
  rename(Genus=genus,`"Species"`=species,`Depth range`=deeps,`Depth zones`=zones)

print_table(crabbo,"genus_depth_table_s","Brachyuran genera exclusive to shallow or mesophotic habitat",hdr_case="none",accuracy=1)
```

\newpage

```{r species-depth-table, results="asis", dependson="taxonomy-validation"}
deep_map <- c(
  "shallow" = str_glue("Shallow ({dr(cc$crabs_shallow_untransformed)} m)"),
  "deep" = "Deep ($\\geq$ 30 m)"
)
crabbo <- cc$crabs %>%
  psmelt() %>%
  mutate(depth_zone = replace(depth_zone,which(depth_zone == "shallow"),str_c(dr(cc$crabs_shallow_untransformed)," m"))) %>%
  group_by(depth_zone) %>%
  filter(Abundance > 0) %>%
  ungroup() %>%
  group_by(species) %>%
  filter(n_distinct(shallow_deep) == 1) %>%
  summarise(zones = str_c(unique(depth_zone),collapse=", "), deeps = str_c(unique(shallow_deep),collapse=", ")) %>%
  ungroup() %>%
  arrange(deeps,zones,species) %>%
  select(deeps,species,zones) %>%
  mutate( deeps = deep_map[deeps] ) %>%
  rename(`"Species"`=species,`Depth range`=deeps,`Depth zones`=zones)

print_table(crabbo,"species_depth_table_s","Brachyuran taxa exclusive to shallow or mesophotic habitat",hdr_case="none",accuracy=1)
```

\newpage

```{r rda-table,results="asis",dependson="rda-analysis"}
print_table(rda_table,"rda_table_s","Significant marginal terms in the dbRDA analysis of crab community variation across the Hawaiian Archipelago",hdr_case="none",accuracy=0.001) 
```

\newpage

```{r nasa-table,results="asis",depenson="init"}
nasa_table <- read_csv(here("data","tables","nasa.csv"))
print_table(nasa_table,"nasa_table_s","NASA data repository filenames for CHL and SST satellite data",accuracy=1,hdr_case=NA,format_numbers=FALSE) 
```

\newpage

## References

<!-- show supplemental bibliography -->
<div class="multi-refs"></div>